---
title: "Initial Data Summaries"
author: "Lee Panter"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```

```{r}
#### Set Working Directory
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/Scripts/Modeling/fbln-cd34"
setwd(WD)


### flowFilter
load("/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/FilteredMergedData/Rdata/flowFilter.RData")

#### mdataFilter
load("/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/FilteredMergedData/Rdata/mdataFilter.RData")

#### seqFilter
load("/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/FilteredMergedData/Rdata/seqFilter.RData")
```


***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce numerical and graphical summaries of relevant models considered in each of the model developed as described in the ReadMe.  

***


<!--  Packages -->  
```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(plyr)
library(nlme)
```


<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
\begin{center}
\LARGE{\underline{\textbf{Begin Script}}}
\end{center}

\vspace{10pt}


```{r}
index.fbln=which(rownames(seqFilter)=="FBLN1")
index.cd34=which(rownames(seqFilter)=="CD34")
fbln=seqFilter[index.fbln,]
cd34=seqFilter[index.cd34,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=mdataFilter$subject.no
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd34, fbln)
```

### Exploratory Data Analysis

Following:

* Quantitative summary tables
* Histogram plots of Predictor and Outcome
* Scatter Plot of Predictor v Outcome
* Numerical five number summaries of Predictor and Outcome


#### Quantitative Data Summaries

```{r}
# variables to include
#  - Subject count
#  - Observation count
#  - Min, Max, Mean, Median, Mode observations

mdataFilterQ=mdataFilter

seqFilterQ=seqFilter
seqFilterQ=t(seqFilterQ)
seqFilterQ=data.frame(seqFilterQ)
seqFilterQ$subject.no=subject.no

mdataSummaryCols=
  mdataFilterQ %>%
    select(subject.no, well, Perc.Mt) %>%
    group_by(subject.no) %>%
    summarise(group_minPerct.Mt=min(Perc.Mt),
              group_maxPerct.Mt=max(Perc.Mt),
              group_avgPerct.Mt=mean(Perc.Mt),
              group_medPerct.Mt=median(Perc.Mt))
mdataSummaryCols
write.csv(mdataSummaryCols, file = "PerctMT2pair.csv")

seqCD34.SummaryCols=
  seqFilterQ %>%
  select(subject.no, CD34, FBLN1) %>% 
  group_by(subject.no) %>% 
  summarise(minCD34 =    min(CD34),
            maxCD34 =    max(CD34),
            avgCD34 =   mean(CD34),
            medCD34 = median(CD34))
seqCD34.SummaryCols

write.csv(seqCD34.SummaryCols, file = "CD34Summaries.csv")


seqFBLN.SummaryCols=
  seqFilterQ %>%
  select(subject.no, CD34, FBLN1) %>% 
  group_by(subject.no) %>% 
  summarise(minFBLN =    min(FBLN1),
            maxFBLN =    max(FBLN1),
            avgFBLN =   mean(FBLN1),
            medFBLN = median(FBLN1))
seqFBLN.SummaryCols

write.csv(seqFBLN.SummaryCols, file = "FBLNSummaries.csv")
```



#### Predictor Summaries
```{r}
# FIVE NUMBER SUMMARY
summary(dat$cd34)


# HISTOGRAM
p1=ggplot(dat, aes(x=cd34,fill=subject.no, color=subject.no))+
  geom_histogram(alpha=0.5, position = "dodge", binwidth = 30)+
  theme(legend.position = "right")
p1
```



#### Outcome Summaries
```{r}
summary(dat$fbln)

# HISTOGRAM
p2=ggplot(dat, aes(x=fbln,fill=subject.no, color=subject.no))+
  geom_histogram(alpha=0.5, position = "dodge", binwidth = 30)+
  theme(legend.position = "right")
p2
```


#### Scatter Plot Outcome ~ Predictor

```{r}
p3=ggplot(dat, aes(x=cd34, y=fbln, color=subject.no))+
  geom_point(alpha=0.5)
p3
```





***

# Transformed Variables (log transformations)

We will apply the transformation $Y=log(x+1)$ to the outcome and response variables to create new-transformed variables.


```{r}
dat$logcd34=log(cd34+1, base = exp(1))
dat$logfbln=log(fbln+1, base = exp(1))


# FIVE NUMBER SUMMARY
summary(dat$logcd34)

# HISTOGRAM
p1=ggplot(dat, aes(x=logcd34,fill=subject.no, color=subject.no))+
  geom_histogram(alpha=0.5, position = "dodge", binwidth = 1)+
  theme(legend.position = "right")
p1

# FIVE NUMBER SUMMARY
summary(dat$logfbln)

# HISTOGRAM
p1=ggplot(dat, aes(x=logfbln,fill=subject.no, color=subject.no))+
  geom_histogram(alpha=0.5, position = "dodge", binwidth = 1)+
  theme(legend.position = "right")
p1


p3=ggplot(dat, aes(x=logcd34, y=logfbln, color=subject.no))+
  geom_point(alpha=0.5)
p3
```
\begin{center}
\LARGE{\underline{\textbf{End Script}}}
\end{center}
<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>


























