---
title: "Linear Model with Fixed Effect Slope"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce, and analyze a linear model with a fixed-effect subject-specific intercept,  and a fixed-effect subject-specific slope, on the finalized data.  It will first produce a single response-predictor model for seq variables, and then analyze the assumptions made for OLS modeling.

***

<!--  ####  Script Dependencies   #### -->


<!--  Packages -->  
```{r, echo=FALSE}
library(ggplot2)
library(plyr)
library(nlme)
library(emmeans)
```


<!--  Working Directory  -->       
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/Scripts/Modeling/fbln-cd34"
setwd(WD)
```


<!-- Data  &  Variables -->       
```{r, echo=FALSE}
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/alldatFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/flowFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/indseqFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/mdataFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/seqFilter.RData")
```
***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
```{r, echo=FALSE}
index.fbln=which(rownames(seqFilter)=="FBLN1")
index.cd34=which(rownames(seqFilter)=="CD34")
fbln=seqFilter[index.fbln,]
cd34=seqFilter[index.cd34,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=mdataFilter$subject.no
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd34, fbln)
dat$logcd34=log(cd34+1, base = exp(1))
dat$logfbln=log(fbln+1, base = exp(1))
```

\vspace{10pt}
# Non-Transformed Model
```{r}
LMwFEslope=lm(fbln~subject.no + cd34:subject.no + cd34 , data = dat)
print(summary(LMwFEslope))
print(AIC(LMwFEslope))

# coefInt=rep(coefficients(LMwFEslope)[1], times=15)
# coefSlp=rep(coefficients(LMwFEslope)[16], times=15)
# 
# coef.int=coefficients(LMwFEslope)[2:15]
# coef.int=c(0,coef.int)
# 
# coef.slopes=coefficients(LMwFEslope)[17:30]
# coef.slopes=c(0,coef.slopes)
# 
# coefInt=coefInt+coef.int
# coefSlp=coefSlp+coef.slopes
# 
# rgLMwFEslope=ref_grid(LMwFEslope)
# lsmeans(rgLMwFEslope, "cd34")
# contrast(rgLMwFEslope, interaction = T)
# 
# 
# plot(fbln~cd34, data=dat, main="Model v Orignal Data")
# for(i in 1:15){
#   abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)
# }

dat$fit=fitted(LMwFEslope)
dat$res=residuals(LMwFEslope)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p


q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q

```

This fitted v residual plot "different" from the first model, but still does not exhibit any symmetrical variation about "0" (or any baseline) which is a VERY strong indicator that residuals are NOT normally distributed.  

# Transformed Model
```{r}
logLMwFEslope=lm(logfbln~subject.no + subject.no:logcd34 + logcd34 , data = dat)
print(summary(logLMwFEslope))
print(AIC(logLMwFEslope))

coefInt=rep(coefficients(logLMwFEslope)[1], times=15)
coefSlp=rep(coefficients(logLMwFEslope)[16], times=15)

coef.int=coefficients(logLMwFEslope)[2:15]
coef.int=c(0,coef.int)

coef.slopes=coefficients(logLMwFEslope)[17:30]
coef.slopes=c(0,coef.slopes)

coefInt=coefInt+coef.int
coefSlp=coefSlp+coef.slopes

plot(logfbln~logcd34, data=dat, main="Model v Orignal Data")
for(i in 1:15){
  xA.temp=is.na(coefInt[i])
  xB.temp=is.na(coefSlp[i])
  if(!xA.temp & !xB.temp)
  {abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)}
}

dat$fit=fitted(logLMwFEslope)
dat$res=residuals(logLMwFEslope)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  theme(legend.position = "none")+
  ylab("Residual Value")
p


q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q

grid.arrange(p,q,nrow=2)
```

<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

