---
title: "Model Summaries"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load("/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/Scripts/Modeling/fbln-cd34/AllCompiledEnviron.RData")
```
\begin{center}
\LARGE{\underline{NON-Tranformed Models}}
\end{center}



# Summary of Intercept and Slope Parameters

## Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-value  \\
\hline
\hline    
M0 (lmod0) & 1.2970 & 1.485e-1 & 8.73 & <2e-16  \\
\hline
M1 (LMwFEint) & 1.8519e1 & 4.067e-1 & 4.554e1 &  <2e-16  \\
\hline
M2 (LMwFEslope) & 1.8414e1 & 4.298e-1 & 4.285e1 & <2e-16 \\
\hline  
M3 (LMMwREint) & 2.75 & 1.37 & 2.01 & 6.4e-2  \\
\hline
M4 (LMMwREslope) & 2.7551 & 1.3588 & 2.03 & 6.2e-2 \\
\hline
M5 (GEE) & 1.424 & 3.34e-1 & 1.819e1** & 2e-5** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)

## Main-Effect Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-Value  \\
\hline
\hline
M0 (lmod0) & 1.4241 & 9.22e-2  & 1.544e1 & <2e-16 \\
\hline
M1 (LMwFEint) & 2.705e-1 & 6.12e-2  & 4.42 & 1.1e-5 \\
\hline
M2 (LMwFEslope) & 3.051e-1 & 7.66e-2  &  3.98 & 7.3e-5 \\
\hline
M3 (LMMwREint) & 2.78e-1  & 6.12e-2 & 4.54 & 6.3e-6  \\
\hline
M4 (LMMwREslope)  & 2.775e-1 &  9.66e-2  & 2.79 & 2.12e-1 \\
\hline
M5 (GEEmod) & 1.297 & 7.18e-1  & 3.26** & 7.1e-2**  \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)



# Model Summaries

## M0 (lmod0)
```{r}
summary(lmod0)
```

## M1 (LMwFEint)
```{r}
summary(LMwFEint)
```

## M2 (LMwFEslope)
```{r}
summary(LMwFEslope)
```

## M3 (LMMwREint)
```{r}
summary(LMMwREint)
```

## M4 (LMMwREslope)
```{r}
summary(LMMwREslope)
```

## M5 (GEEmod)
```{r}
summary(GEEmod)
```

# Nested Model Comparisons
```{r}
# M0 < M1
anova(lmod0, LMwFEint )

# M1 < M2
anova(LMwFEint, LMwFEslope)

# M3 < M4
anova(LMMwREint, LMMwREslope)
```

# AIC Measurement Statistics
```{r}
AIC(lmod0,LMwFEint, LMwFEslope, LMMwREint, LMMwREslope)
```
***

\begin{center}
\LARGE{\underline{Tranformed Models}}
\end{center}

# Summary of Intercept and Slope Parameters

## Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-Value \\
\hline
\hline
M0 (loglmod0) & 3.510e-1 & 2.45e-2 & 1.43e1 & < 2e-16 \\
\hline
M1 (logLMwFEint) & 2.7572 & 6.52e-2 & 4.23e1 & < 2e-16 \\
\hline
M2 (logLMwFEslope) & 2.7973 & 7.68e-2 & 3.643e1 & < 2e-16 \\
\hline
M3 (logLMMwREint) & 6.53e-1 & 2.22e-1 & 2.94 & 1.1e-2 \\
\hline
M4 (logLMMwREslope) & 6.491e-1 & 2.223e-1 & 2.92 & 1.1e-2 \\
\hline
M5 (logGEEmod) & 3.51e-1 & 1.25e-1 & 7.85** & 5.09e-3** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)


## Main-Effect Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-value \\
\hline
\hline
M0 (loglmod0) & 7.884e-1 & 4.92e-2  & 1.6e+1 & < 2e-16 \\
\hline
M1 (logLMwFEint) & 1.306e-1 & 3.42e-2  & 3.82 & 1.4e-4 \\
\hline
M2 (logLMwFEslope) & 8.38e-2 & 5.89e-2  &  1.42 & 1.5492e-1 \\
\hline
M3 (logLMMwREint) & 1.35e-1 & 3.42e-2  & 3.95 & 8.4e-5 \\
\hline
M4 (logLMMwREslope) & 1.705e-1  & 7.29e-2 & 2.34  & 6.7e-2 \\
\hline
M5 (logGEEmod) & 7.88e-1 &  2.2e-1  & 1.281e+1** & 3.4e-4** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)



# Model Summaries

## M0 (loglmod0)
```{r}
summary(loglmod0)
```

## M1 (logLMwFEint)
```{r}
summary(logLMwFEint)
```

## M2 (logLMwFEslope)
```{r}
summary(logLMwFEslope)
```

## M3 (logLMMwREint)
```{r}
summary(logLMMwREint)
```

## M4 (logLMMwREslope)
```{r}
summary(logLMMwREslope)
```

## M5 (logGEEmod)
```{r}
summary(logGEEmod)
```

# Nested Model Comparisons
```{r}
# M0 < M1
anova(loglmod0, logLMwFEint )

# M1 < M2
anova(logLMwFEint, logLMwFEslope)

# M3 < M4
anova(logLMMwREint, logLMMwREslope)
```


# AIC Measurement Statistics
```{r}
AIC(loglmod0, logLMwFEint, logLMwFEslope, 
    logLMMwREint, logLMMwREslope)
```














<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

