---
title: "Model Descriptions"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---
***
<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```

<!--  ####  Description   #### -->

# Description:
This script will outline the intended model fits contained within the "Modeling" folder.  It will provide the theoretical foundation for each model, including:

* Parameters
* Coeficients
* Distributions, and distributional assumptions
* Additional assumptions made while modeling

Each model will have multiple sections.  Each sections will pertain to a separate model, or other distinction (e.g. alternative assumptions). 

Nomenclature, and notation is outlined first, and this should be universally applicable across all models.

***

<!--  ####  Script Dependencies   #### -->

<!--  Packages -->  

<!--  Working Directory  -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/ModelDescription"
setwd(WD)
```

<!------------------------------------------------------------------------------>





<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>

***

# Nomenclature, Notation and Definitions

## Models

The following sections will represent the five models to be fit:

1. Linear Model with Fixed Effect Intercept (LMwFEint)
2. Linear Model with Fiexed Effect Slope (LMwFEslope)
3. Linear Mixed Model with Random Intercept (LMMwREint)
4. Linear Mixed Model with Random Slope (LMMwREslope)
5. Generalized Estimating Equations (GEE)

## Variable Sets
\begin{align*}
\Omega_{SEQ} &= \left \{SEQ_{p}: \ p=1, \ldots, 38354    \right \}\\
&= \left \{ SEQ_{1}=``A1BG-AS1", \ SEQ_{2}= ``A1CF", \ \ldots, \  SEQ_{38353}=``SNOZ5", \ SEQ_{38354}=``SNOSNR66"  \right \}\\
\\
\Omega_{FLOW} &= \left \{ FLOW_{p}: \ p=1, \ldots, 19    \right \}\\
&= \left \{ FLOW_{1}=``FSC.A", \ FLOW_{2}= ``FSC.W", \ \ldots, \  FLOW_{18}=``CD27", \ FLOW_{19}=``CD235a"  \right \}\\
\\
\Omega_{META} &= \left \{ META_{p}: \ p=1, \ldots, 14    \right \}\\
&= \left \{ META_{1}=``measurement.name", \ \ldots, \ META_{19}=``Perc.Mt"  \right \}\\
\end{align*}

We choose:

\begin{center}
$p_{resp}^{seq} \in \Omega_{seq}$\\
\vspace{5pt}
and\\
\vspace{5pt}
$p_{resp}^{flow} \in \Omega_{flow}$\\
\end{center}

to model as response values, using the values:

\begin{center}
$p_{cov}^{seq} \in \Omega_{seq} \backslash \left \{p_{resp}^{seq}\right \}$\\
\vspace{5pt}
and\\
\vspace{5pt}
$p_{cov}^{flow} \in \Omega_{flow}\backslash \left \{p_{resp}^{flow}\right \}$\\
\end{center}

as covariates.

We also let 

* $i = 1, \ldots, N$ Represent subject
* $j = 1, \ldots, n_{i}$ Represent measurement within subject
* $v \in \left \{ seq, flow \right \}$

***

# Model 0: OLS Referrence Model

$$\log \left [ \left( Y_{p_{resp}^{v}}\right)_{ij} \right ]  = \beta_{0} + \beta_{1} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] + \epsilon_{ij} $$

where 
$$\epsilon_{ij} \sim N\left(\mathbf{0}, \sigma_{\epsilon}^{2} \mathbf{I}_{n_{i}}   \right) $$

# Model 1: Linear Model with Fixed Effect Intercept (LMwFEint)

$$\log \left [ \left( Y_{p_{resp}^{v}}\right)_{ij} \right ]  = \beta_{0} + \beta_{1i} \left( subject \right)_{i} +   \beta_{2} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] + \epsilon_{ij} $$

where 
$$\epsilon_{ij} \sim N\left(\mathbf{0}, \sigma_{\epsilon}^{2} \mathbf{I}_{n_{i}}   \right) $$


# Model 2: Linear Model with Fixed Effect Slope (LMwFEslope)

$$\log \left [ \left( Y_{p_{resp}^{v}}\right)_{ij} \right ]  = \beta_{0} + \beta_{1i} \left( subject \right)_{i} +   \beta_{2i} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] * \left(subject \right)_{i}  + \epsilon_{ij} $$

again with:

$$\epsilon_{ij} \sim N\left(\mathbf{0}, \sigma_{\epsilon}^{2} \mathbf{I}_{n_{i}}   \right) $$

# Model 3: Linear Mixed Model with Random Intercept (LMMwREint)

$$\log \left [ \left( Y_{p_{resp}^{v}}\right)_{ij} \right ]  = \beta_{0} + \beta_{1} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] + b_{0i} \left(subject \right)_{i}  + \epsilon_{ij} $$

where 

$$b_{0i} \sim N\left(0 , \sigma_{b}^{2}    \right) $$

$$\epsilon_{ij} \sim N\left(\mathbf{0}, \sigma_{\epsilon}^{2} \mathbf{I}_{n_{i}}   \right) $$



# Model 4: Linear Mixed Model with Random Slope (LMMwREslope)


\begin{align*}
\log \left [ \left( Y_{p_{resp}^{v}}\right)_{ij} \right ]  &= \beta_{0} + \beta_{1} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] +  \\
&= b_{0i} \left(subject \right)_{i} + b_{1i}\log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ]*\left( subject \right)_{i}    + \epsilon_{ij} \\
\end{align*}

where 
$$
\mathbf{b} = 
\begin{bmatrix}
b_{0i} \\
b_{1i} \\
\end{bmatrix} 
\sim N(\mathbf{0}, \mathbf{G})
$$

$$
\mathbf{G}=
\begin{bmatrix}
\sigma_{b}^{2} & 0 \\
0 & \sigma_{b}^{2} \\
\end{bmatrix}
$$



$$\epsilon_{ij} \sim N\left(\mathbf{0}, \sigma_{\epsilon}^{2} \mathbf{I}_{n_{i}}   \right) $$


# Model 5: Generalized Estimating Equations (GEE)

The GEE fitting process is a semi-parametric method that uses intial estimates obtained from parametric methods of estimating Generalized Linear Models, to iteratively re-estimate parameters according to a working covariance matrix that appropriately models within-subject variability until a desired threshold of convergence is met.  

Accordingly, the following criteria are needed to fully specify a GEE estimate:

* An exponential family to specify error structure in initial conditions, with a similarly defined link function
* A linear predictor to specify covariate information
* A working covariance structure

In the case of the model being fit here, we assume that:

* For initial conditions: $Y_{ij} \sim N \left(\mu, \sigma^{2}\right)$ with link function $g(x)=x$
* \textit{Alternate}--For initial conditions: $Y_{ij} \sim Poisson( \lambda)$ with link function $g(x)=log(x)$

\newpage

* We will use the linear predictor:

$$\eta_{ij} = \beta_{0} + \beta_{1i} \left( subject \right)_{i} +   \beta_{2i} \log \left [ \left(X_{p_{cov}^{v}}\right)_{ij} \right ] * \left(subject \right)_{i} $$
* We will be assuming that observations follow an independent covariance structure, which implies that:

$$
corr_{i}=R_{i}=
\begin{bmatrix}
1 & 0 & 0 & \cdots & 0 \\
0 & 1 & 0 & \cdots & 0 \\
0 & 0 & 1 & \cdots & 0 \\
\vdots & \vdots &  \vdots & \ddots  & \vdots \\
0 & 0 & 0 & \cdots & 1 \\
\end{bmatrix}_{\left(n_{i} \ x \ n_{i} \right) }
$$




































<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

