---
csl: mathematical-biosciences-and-engineering.csl
output:
  pdf_document: 
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
  word_document: 
geometry: margin=1.0in
fontsize: 12pt
subtitle: Version 2.2-Results
bibliography: Bib_AccRefSheet.bib
---


<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/V2.2/5 Results"
setwd(WD)
```
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
# Results

- The table below (\textbf{Table} \hspace{5pt}) shows the estimate, standard error, test statistic, and p-values associated with the fixed-effect slope parameter as estimated by each modeling method for the $MALAT1 \sim CD19$ variable pairing.

\vspace{30pt}

\small{
\begin{center}
\centering
\begin{tabular}{|m{0.15\textwidth}|m{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  4.918e-2 &  1.455e-2  & 3.381* &  7.47e-4\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  4.833e-2 &  1.381e-2 & 3.500* &  4.84e-4\\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  4.920e-2 &  1.374e-2 & 3.579*  &  3.6e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 5.938e-2 &  3.538e-2  & 1.678* &  1.19e-1 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 4.918e-2 & 1.455e-2 & 3.381** & 7.47e-4 \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt} : Summary table for each model method, $MALAT1 \sim CD19$ variable pairing. ** Approximate Wald-Z distribution
\end{center}
}

\newpage

- The table below (\textbf{Table} \hspace{5pt}) shows the estimate, standard error, test statistic, and p-values associated with the fixed-effect slope parameter as estimated by each modeling method for the $FBLN1 \sim CD34$ variable pairing.

\vspace{30pt}

\small{
\begin{center}
\centering
\begin{tabular}{|m{0.15\textwidth}|m{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  7.884e-1 &  4.92e-2  & 4.002* & <2e-16\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  1.31e-1 & 3.42e-2  & 3.818* & 1.42e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  1.35e-1 & 3.42e-2  & 3.95* & 8.4e-5 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 1.705e-1  & 7.29e-2 & 2.34*  & 6.7e-2 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 7.884e-1 & 4.92e-2 & 4.002** & < 2e-16 \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt} : Summary table for each model method, $FBLN1 \sim CD34$ variable pairing. ** Approximate Wald-Z distribution
\end{center}
}




- The fixed effect slope parameter is being considered for comparison because we are interested in quantifying correlation between single-cell values at the subject level.

- Displayed in \textbf{table \hspace{5pt}}, \textbf{table \hspace{5pt}}, and \textbf{table \hspace{5pt}} below are percent change in: parameter estimate, standard error, and test statistic (respectively) for the $MALAT1 \sim CD19$variable pairing as defined by:

$$\text{Percent Change} \ [A]_{ij} = \left(\frac{A_{j}-A_{i}}{A_{i}}\right) * 100$$
\newpage
\begin{center}
\underline{\textbf{Parameter Estimate Percent Change}}\\
\vspace{5pt}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0 &  -1.7283 &  0.0407 & 20.7401 &  0.0000 \\
\hline
LM-FE & 1.7587 &  0 &  1.8001 & 22.8636  & 1.7587  \\      
\hline
LMM-RI & -0.0407 & -1.7683 &  0 & 20.6911 & -0.0407  \\
\hline
LMM-RS & -17.1775 & -18.6090 & -17.1438 & 0 & -17.1775 \\
\hline
GEE & 0.0000 & -1.7283  & 0.0407 & 20.7401  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: Main effect slope percent change matrix, $MALAT1 \sim CD19$ variable pairing
\end{center}

\vspace{10pt}


\begin{center}
\underline{\textbf{Standard Error Percent Change}}\\
\vspace{5pt}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -5.0859  & -5.5670   & 143.1615  & 0.0000  \\
\hline
LM-FE & 5.3584   & 0   & -0.5069   & 156.1912  & 5.3584  \\     
\hline
LMM-RI & 5.8952   & 0.5095   & 0    & 157.4964  & 5.8952  \\
\hline
LMM-RS & -58.8751 & -60.9666 & -61.1645  & 0    & -58.8751  \\
\hline
GEE &  0.0000   & -5.0859  & -5.5670   & 143.1615  & 0        \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: Main effect slope standard error percent change matrix, $MALAT1 \sim CD19$ variable pairing
\end{center}


\vspace{10pt}

\begin{center}
\underline{\textbf{Test Statistic Percent Change}}\\
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & 3.5197   & 5.8563    & -50.3697  & 0.0000 \\
\hline
LM-FE & -3.4000  & 0   & 2.2571    & -52.0571  & -3.4000 \\
\hline
LMM-RI & -5.5323  & -2.2073  & 0    & -53.1154  & -5.5323  \\
\hline
LM-RS & 101.4899 & 108.5816 & 113.2896  & 0    & 101.4899 \\
\hline
GEE & 0.0000  & 3.5197   & 5.8563    & -50.3697  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: Main effect slope test statistic percent change matrix, $MALAT1 \sim CD19$ variable pairing
\end{center}


- Displayed in \textbf{table \hspace{5pt}}, \textbf{table \hspace{5pt}}, and \textbf{table \hspace{5pt}} below are percent change in: parameter estimate, standard error, and test statistic (respectively) for the $FBLN1 \sim CD34$ variable pairing as defined by:

$$\text{Percent Change} \ [A]_{ij} = \left(\frac{A_{j}-A_{i}}{A_{i}}\right) * 100$$

\vspace{10pt}

\begin{center}
\underline{\textbf{Parameter Estimate Percent Change}}\\
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -83.3841 & -82.8767  & -78.3739 & 0.0000  \\
\hline
LM-FE & 501.8321 & 0   & 3.0534    & 30.1527  & 501.8321 \\
\hline
LM-RI & 484.0000 & -2.9630  & 0    & 26.2963  & 484.0000  \\
\hline
LM-RS & 362.4047 & -23.1672 & -20.8211  & 0   & 362.4047  \\
\hline
GEE & 0.0000   & -83.3841 & -82.8767  & -78.3739 & 0  \\
\hline
\end{tabular}

\vspace{10pt}

\textbf{Table} \hspace{5pt}: Main effect slope percent change matrix, $FBLN1 \sim CD34$ variable pairing
\end{center}

\vspace{10pt}

\begin{center}
\underline{\textbf{Standard Error Percent Change}}\\
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -30.4878 & -30.4878  & 48.1707   & 0.0000  \\
\hline
LM-FE & 43.8596  & 0   & 0.0000    & 113.1579  & 43.8596  \\
\hline
LM-RI & 43.8596  & 0.0000   & 0    & 113.1579  & 43.8596  \\
\hline
LM-RS & -32.5103 & -53.0864 & -53.0864  & 0    & -32.5103  \\
\hline
GEE &  0.0000   & -30.4878 & -30.4878  & 48.1707   & 0  \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: Main effect slope standard error percent change matrix, $FBLN1 \sim CD34$ variable pairing
\end{center}

\newpage

\begin{center}
\underline{\textbf{Test Statistic Percent Change}}\\
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & -4.5977  & -1.2994   & -41.5292  & 0.0000  \\
\hline
LM-FE & 4.8193  & 0   & 3.4573    & -38.7114  & 4.8193  \\
\hline
LM-RI & 1.3165  & -3.3418  & 0    & -40.7595  & 1.3165  \\
\hline
LM-RS & 71.0256 & 63.1624  & 68.8034   & 0    & 71.0256  \\
\hline
GEE & 0.0000  & -4.5977  & -1.2994   & -41.5292  & 0  \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: Main effect slope test statistic percent change matrix, $FBLN1 \sim CD34$ variable pairing
\end{center}

## Parameter Estimates

- Models LM and GEE have similar estimates
- Models LM-FE, LM-RI, and LM-RS are also all similar 
- Estimates are less similar when comparing between the two clusters of similarity mentionioned
- Accounting for subject-specific outcomes results in a different estimate than when modeling population-averages, i.e. there is an effect from subject-level correlation
- The effect is more pronounced in the FBLN1~CD34 pairing than the MALAT1~CD19 pairing



## Standard Error Estimates

- The standard error of the fixed effect slope parameter estimate is highest for the random slope model
- The SE is lowest for the random intercept model
- The fixed-effect subject-specific intercept had a lower SE than either of the population-average parameter interpretaion models LM and GEE.
    





## Test Statistics

- Changes in test statistics for the fixed-effect main-effect slope were smaller between LM and GEE, as well as between LM-FE, LMM-RI, and LMM-RS with few exceptions.
- There were very small differences between LM and GEE
- Changes to the LMM-RS resulted in calculated p-value increases with two and three orders of magnitude.
    
<!------------------------------------------------------------------------------>

## Nested Model Comparisons

### Test for FE Subject Specific Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & Resid DF & RSS & DF & Sum of Squares & F-stat & P(>F) \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LM & 1108 & 1167.76 & & & & \\
 & LM-FE & 1094 & 935.89 & 14 & 231.87 & 19.36 & < 2.2e-16 \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LM & 1108 & 650.51 & & & & \\
 & LM-FE & 1094 & 214.92 & 14 & 435.59 & 158.38 & < 2.2e-16 \\
\hline
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: ANOVA nested model comparison table for testing the inclusion of the subject-specific fixed-effect intercept
\end{center}


\textbf{Table} \hspace{5pt} above is a nested model comparison, the result of which is an F-test statistic telling us that there is very strong evidence to support the inclusion of the subject-specific fixed-effect intercept into the LM model. 




### Test for RE Subject Specific Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & df & AIC & logLik & L.Ratio & p-value \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LM & 3 & 3224.097 & -1609.048 &  & \\
 & LMM-RI & 4 & 3032.024 & -1512.012 & 194.0722 & <1e-4  \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LM & 3 & 2572.807 & -1283.403 &  & \\
 & LMM-RI & 4 & 1438.086 & -715.043 & 1136.72 & <1e-4  \\
\hline
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: ANOVA nested model comparison table for testing the inclusion of the subject-specific random effect intercept
\end{center}


\textbf{Table} \hspace{5pt} above is a nested model comparison, the result of which is a likelihood ratio statistic telling us that there is very strong evidence to support the inclusion of the subject-specific random effect intercept into the LM model. 



### Test for RE Subject Specific Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & df & AIC & logLik & L.Ratio & p-value \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LMM-RI & 4 & 3032.024 & -1512.012 &  & \\
 & LMM-RS & 6 & 2993.820 & -1490.910 & 42.20503 & <1e-4  \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LMM-RI & 4 & 1438.086 & -715.043 &  & \\
 & LMM-RS & 6 & 1438.068 & -713.034 & 4.018095 & 0.1341  \\
\hline
\end{tabular}

\vspace{5pt}

\textbf{Table} \hspace{5pt}: ANOVA nested model comparison table for testing the inclusion of the subject-specific random effect slope
\end{center}


\textbf{Table} \hspace{5pt} above is a nested model comparison, the result of which is a likelihood ratio statistic telling us that there is very strong evidence to support the inclusion of the subject-specific random effect slope into the LMM-RI model for the MALAT1~CD19 variable pairing. However, there is insufficient evidence to support the inclusion of the subject-specific random effect slope into the LMM-RI model for the FBLN~CD34 variable pairing.





# References

\bibliography{Bib_AccRefSheet}

