---
title: "Final Result Comparisons MALAT1 vs CD19"
author: "Lee Panter"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
bibliography: BibdotBib.bib
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE,echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will compile all relevant data and model summary information for the MALAT1 ~ CD19 pairings.  

***

<!--  ####  Script Dependencies   #### -->
\begin{center}
\LARGE{\underline{\textbf{Script Dependencies}}}
\end{center}


<!--  Packages -->  
## Package Dependencies

```{r}
library(ggplot2)
library(plyr)
library(nlme)
```


<!--  Working Directory  -->       
## SetWorking Directory & Seed

```{r}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/Scripts/ResultSummaries"
setwd(WD)
set.seed(123)
```



<!-- Data  &  Variables -->       

***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
\begin{center}
\LARGE{\underline{\textbf{Percent Change Matrices}}}
\end{center}

```{r, echo=FALSE}
perct.change=function(old, new){
  out=(new-old)/old
  return(out)
}

perct.change.matrix=function(vectin){
  l=length(vectin)
  mat.out=matrix(NA, nrow=l, ncol=l)
  for(i in 1:l){
    for(j in 1:l){
      mat.out[i,j]=perct.change(vectin[i], vectin[j])
    }
  }
  return(mat.out)  
}


int.estimate=c((7.8294)*10^(3),
               (10.1125)*10^(3),
               (9.559)*10^(3),
               (7.3994)*10^(3),
               (7.4050)*10^(3),
               (7.8294)*10^(3))


int.se=c((2.348)*10^(2),
         (8.4495)*10^(2), 
         (8.956)*10^(2),
         (7.6899)*10^(2),
         (7.677)*10^(2),
         (9.8883)*10^(2))



slope.estimate=c(0.7132,
                   2.562,
                   (1.765)*10^(1),
                   2.495,
                   2.168,
                   0.7132)


slope.se=c(1.5426,
           1.501,
           8.246,
           1.491,
           1.797,
           2.5028)


log.int.estimate=c(8.46182,
                   7.54861,
                   6.97932,
                   8.41374,
                   8.39722,
                   8.4618)


log.int.se=c(0.04568,
             0.12291,
             0.13896,
             0.11825,
             0.13957,
             0.1842)


log.slope.estimate=c(0.04918,
                     0.04833,
                     0.51428,
                     0.04920,
                     0.05938,
                     0.0492)

log.slope.se=c(0.01455,
               0.01381,
               0.06017,
               0.01374,
               0.03538,
               0.0397)
```

# Untransformed Percent Change Matrices

## Intercept 

### Estimate
```{r}
round(perct.change.matrix(int.estimate), 3)
```

### Standard Error
```{r}
round(perct.change.matrix(int.se), 3)
```

## Slope

### Estimate
```{r}
round(perct.change.matrix(slope.estimate), 3)
```

### Standard Error
```{r}
round(perct.change.matrix(slope.se), 3)
```

\newpage

# Log-Transformed Models

## Intercept 

### Estimate
```{r}
round(perct.change.matrix(log.int.estimate), 3)
```

### Standard Error
```{r}
round(perct.change.matrix(log.int.se), 3)
```

## Slope

### Estimate
```{r}
round(perct.change.matrix(log.slope.estimate), 3)
```

\newpage

### Standard Error
```{r}
round(perct.change.matrix(log.slope.se), 3)
```


<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

