---
csl: mathematical-biosciences-and-engineering.csl
output:
  pdf_document: 
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=1.0in
fontsize: 12pt
bibliography: Bib_AccRefSheet.bib
---


<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/vFINAL/7 Appendix and Post Script"
setwd(WD)
```
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>

# Appendix

### Table A1

\begin{center}

\textbf{\large{CD19 Summaries}}

\begin{tabular}{|c|c|c|c|c|}
\hline 
Subject Number & Minimum & Maximum & Average & Median \\ 
\hline
\hline
5 &  0 & 678 & 36.6724 & 0.0 \\ 
\hline  
6 &  0 & 299 & 36.6860 & 7.5 \\ 
\hline  
7 &  0 & 10 & 2.1250 & 1.0 \\ 
\hline  
9 &  0 & 1052 & 89.4194 & 3.0 \\ 
\hline  
10 & 0 & 158 & 37.5714 & 2.0 \\ 
\hline   
11 & 0 & 339 & 28.3178 & 1.0 \\ 
\hline   
13 & 0 & 629 & 56.0841 & 18.0 \\ 
\hline   
14 & 0 & 251 & 42.2600 & 19.0 \\ 
\hline   
15 & 0 & 148 & 26.6000 & 0.0 \\ 
\hline   
17 & 0 & 982 & 112.3770 & 16.0 \\ 
\hline   
19 & 0 & 665 & 59.3386 & 5.0 \\ 
\hline   
20 & 0 & 287 & 40.1200 & 23.0 \\ 
\hline   
22 & 0 & 380 & 43.4483 & 1.0 \\ 
\hline   
24 & 0 & 282 & 55.0127 & 27.0 \\ 
\hline  
26 & 0 & 1624 & 268.4151 & 110.0 \\  
\hline
\end{tabular}

\vspace{5pt}

Table A1: Predictor $CD19$ variable summaries ($CD19 \sim MALAT1$)
\end{center}


\newpage

### Table A2


\begin{center}

\textbf{\large{MALAT1 Summaries}}

\begin{tabular}{|c|c|c|c|c|}
\hline 
Subject Number & Minimum & Maximum & Average & Median \\ 
\hline
\hline
5  & 67 & 40812 & 10206.3621 & 9195.0 \\ 
\hline 
6  & 757 & 30774 & 11568.2791 & 11689.0 \\ 
\hline 
7  & 441 & 17916 & 6868 & 4039.5 \\ 
\hline 
9  & 311 & 18239 & 5703.9355 & 5983.0 \\ 
\hline 
10 & 1875 & 17160 & 6638.5714 & 6190.0 \\ 
\hline
11 & 349 & 34082 & 9716.0280 & 8826.0 \\ 
\hline 
13 & 99 & 25572 & 5867.9439 & 4895.0 \\ 
\hline 
14 & 355 & 15740 & 6154.1500 & 5720.5 \\ 
\hline 
15 & 157 & 11923 & 3839.0800 & 3467.0 \\ 
\hline 
17 & 337 & 8342 & 2960.2541 & 2692.0 \\ 
\hline 
19 & 227 & 91961 & 13959.9843 & 10125.0 \\ 
\hline 
20 & 379 & 21736 & 7301.4133 & 6417.0 \\ 
\hline 
22 & 161 & 28429 & 6881.7471 & 5068.0 \\ 
\hline 
24 & 240 & 42792 & 6248.8228 & 5955.0 \\ 
\hline 
26 & 1114 & 32426 & 8463.1698 & 6426.0 \\ 
\hline  
\end{tabular}

\vspace{5pt}

Table A2: Response $MALAT1$ variable summaries ($CD19 \sim MALAT1$)

\end{center}


\newpage


### Table A3

\begin{center}

\textbf{\large{CD34 Summaries}}


\begin{tabular}{|c|c|c|c|c|}
\hline 
Subject Number & Minimum & Maximum & Average & Median \\ 
\hline
\hline
5  & 0 & 19 & 3.0517 & 1 \\ 
\hline 
6  & 0 & 0 & 0 & 0 \\ 
\hline 
7  & 0 & 0 & 2 & 1 \\ 
\hline 
9  & 0 & 6 & 0.4516 & 0 \\ 
\hline 
10 & 0 & 5 & 0.6667 & 0 \\ 
\hline 
11 & 0 & 7 & 1.2056 & 1 \\ 
\hline 
13 & 0 & 0 & 0 & 0 \\ 
\hline 
14 & 0 & 1 & 0.4000 & 0 \\ 
\hline 
15 & 0 & 0 & 0 & 0 \\ 
\hline 
17 & 0 & 0 & 0 & 0 \\ 
\hline 
19 & 0 & 0 & 0 & 0 \\ 
\hline 
20 & 0 & 2 & 0.1867 & 0 \\ 
\hline 
22 & 0 & 4 & 0.3563 & 0 \\ 
\hline 
24 & 0 & 5 & 0.2911 & 0 \\ 
\hline 
26 & 0 & 0 & 0 & 0 \\  
\hline
\end{tabular}

\vspace{5pt}

Table A3: Predictor $CD34$ variable summaries ($CD34 \sim FBLN1$)

\end{center}



\newpage

### Table A4

\begin{center}

\textbf{\large{FBLN1 Summaries}}


\begin{tabular}{|c|c|c|c|c|}
\hline 
Subject Number & Minimum & Maximum & Average & Median \\ 
\hline
\hline
5  & 3 & 41 & 19.3448 & 18 \\ 
\hline 
6  & 0 & 0 & 0 & 0 \\ 
\hline 
7  & 0 & 16 & 4.2500 & 3 \\ 
\hline 
9  & 0 & 8 & 1.8710 & 1 \\ 
\hline 
10 & 0 & 30 & 11.9524 & 10 \\ 
\hline 
11 & 0 & 8 & 1.5140 & 1 \\ 
\hline 
13 & 0 & 1 & 0.0093 & 0 \\ 
\hline 
14 & 0 & 5 & 0.5700 & 0 \\ 
\hline 
15 & 0 & 1 & 0.0400 & 0 \\ 
\hline 
17 & 0 & 3 & 0.0246 & 0 \\ 
\hline 
19 & 0 & 2 & 0.0157 & 0 \\ 
\hline 
20 & 0 & 9 & 2.5867 & 2 \\ 
\hline 
22 & 0 & 11 & 0.9885 & 0 \\ 
\hline 
24 & 0 & 4 & 0.4557 & 0 \\ 
\hline 
26 & 0 & 0 & 0 & 0 \\ 
\hline
\end{tabular}

\vspace{5pt}

Table A4: Response $FBLN1$ variable summaries ($CD34 \sim FBLN1$)
\end{center}


<!------------------------------------------------------------------------------>

\newpage

# Code and Data

All code for the above analysis was written and evaluated in RStudio Version 1.2.1335, and is available for download at the following GitHub repository: 

https://github.com/leepanter/MSproject_RBC.git

Additionally, a link to all necessarry and referrence data files (including original data) are contained in the following Google Drive:

https://drive.google.com/open?id=1gjHaMJG0Y_kPYWj5bIE4gRJU5z9R2Wqb


<!------------------------------------------------------------------------------>



# References

\bibliography{Bib_AccRefSheet}


<!------------------------------------------------------------------------------>

