---
title: "Initial Data Summaries"
author: "Lee Panter"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce numerical and graphical summaries of relevant models considered in each of the model developed as described in the ReadMe.  

***

<!--  Packages -->  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plyr)
library(nlme)
library(dplyr)
```


<!--  Working Directory  -->       
```{r, echo=FALSE}
WD="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/InitialDataSummaries"
setwd(WD)
```


<!-- Data  &  Variables -->       
```{r, echo=FALSE}
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/alldatFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/flowFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/indseqFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/mdataFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/seqFilter.RData")
```
***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
\begin{center}
\LARGE{\underline{\textbf{Begin Script}}}
\end{center}

\vspace{10pt}

```{r, echo=FALSE}
index.malat1=which(rownames(seqFilter)=="MALAT1")
index.cd19=which(rownames(seqFilter)=="CD19")
mala=seqFilter[index.malat1,]
cd19=seqFilter[index.cd19,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=mdataFilter$subject.no
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd19, mala)
```


### Exploratory Data Analysis

Following:

* Quantitative summary tables
* Histogram plots of Predictor and Outcome
* Scatter Plot of Predictor v Outcome
* Numerical five number summaries of Predictor and Outcome


#### Quantitative Data Summaries

```{r}
# variables to include
#  - Subject count
#  - Observation count
#  - Min, Max, Mean, Median, Mode observations
#  - 

mdataFilterQ=mdataFilter

seqFilterQ=seqFilter
seqFilterQ=t(seqFilterQ)
seqFilterQ=data.frame(seqFilterQ)
seqFilterQ$subject.no=subject.no

mdataCountCols=
    mdataFilterQ %>%
      select(subject.no, well, Perc.Mt) %>% 
      group_by(subject.no) %>% 
      count(subject.no)
mdataCountCols

write.csv(mdataCountCols, file = "SubjectCounts.csv")


mdataSummaryCols=
  mdataFilterQ %>%
    select(subject.no, well, Perc.Mt) %>% 
    group_by(subject.no) %>% 
    summarise(group_minPerct.Mt=min(Perc.Mt),
              group_maxPerct.Mt=max(Perc.Mt),
              group_avgPerct.Mt=mean(Perc.Mt),
              group_medPerct.Mt=median(Perc.Mt))

write.csv(mdataSummaryCols, file = "SubjectPerctMT.csv")

seqCD19.SummaryCols=
  seqFilterQ %>%
  select(subject.no, CD19, MALAT1) %>% 
  group_by(subject.no) %>% 
  summarise(group_minCD19 =    min(CD19),
            group_maxCD19 =    max(CD19),
            group_avgCD19 =   mean(CD19),
            group_medCD19 = median(CD19))
seqCD19.SummaryCols

write.csv(seqCD19.SummaryCols, file = "CD19Summaries.csv")


seqMALA.SummaryCols=
  seqFilterQ %>%
  select(subject.no, CD19, MALAT1) %>% 
  group_by(subject.no) %>% 
  summarise(group_minMALA =    min(MALAT1),
            group_maxMALA =    max(MALAT1),
            group_avgMALA =   mean(MALAT1),
            group_medMALA = median(MALAT1))
seqMALA.SummaryCols

write.csv(seqMALA.SummaryCols, file="MALASummaries.csv")            
```




#### Predictor Summaries
```{r}
# FIVE NUMBER SUMMARY
summary(dat$cd19)

# HISTOGRAM
# p1=ggplot(dat, aes(x=cd19,fill=subject.no, color=subject.no))+
#   geom_histogram(alpha=0.5, position = "dodge", binwidth = 30)+
#   theme(legend.position = "right")
# p1

hist(cd19, xlab = "cd19", breaks= 100 )
```

#### Outcome Summaries
```{r}
# FIVE NUMBER SUMMARY
summary(dat$mala)

# HISTOGRAM
# p2=ggplot(dat, aes(x=mala,fill=subject.no, color=subject.no))+
#   geom_histogram(alpha=0.5, position = "dodge", binwidth = 30)+
#   theme(legend.position = "right")
# p2

hist(mala, xlab = "MALAT1", breaks = 100)
```

#### Scatter Plot Outcome ~ Predictor

```{r}
p3=ggplot(dat, aes(x=cd19, y=mala, color=subject.no))+
  geom_point(alpha=0.5)
p3
```




***

# Transformed Variables (log transformations)

We will apply the transformation $Y=log(x+1)$ to the outcome and response variables to create new-transformed variables.


```{r}
dat$logcd19=log(cd19+1, base = exp(1))
dat$logmala=log(mala+1, base = exp(1))
logcd19=dat$logcd19
logmala=dat$logmala

# FIVE NUMBER SUMMARY
summary(dat$logcd19)

# HISTOGRAM
# p1=ggplot(dat, aes(x=logcd19,fill=subject.no, color=subject.no))+
#   geom_histogram(alpha=0.5, position = "dodge", binwidth = 1)+
#   theme(legend.position = "right")
# p1

hist(logcd19, xlab = "logcd19", breaks = 100)

# FIVE NUMBER SUMMARY
summary(dat$logmala)

# HISTOGRAM
# p1=ggplot(dat, aes(x=logmala,fill=subject.no, color=subject.no))+
#   geom_histogram(alpha=0.5, position = "dodge", binwidth = 1)+
#   theme(legend.position = "right")
# p1

hist(logmala, xlab = "logmala", breaks = 100)

p3=ggplot(dat, aes(x=logcd19, y=logmala, color=subject.no))+
  geom_point(alpha=0.5)
p3
```
\begin{center}
\LARGE{\underline{\textbf{End Script}}}
\end{center}
<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

