---
title: "Linear Mixed Model with Random Effect Intercept"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce, and analyze a linear mixed model with the following components evaluated on the final data:

* a subject-specific, fixed effect intercept
* a subject-specific, fixed effect slope
* a subject-specific, random effect intercept

It will first produce a single response-predictor model for seq variables, and then analyze the assumptions made for LMM modeling.

***

<!--  ####  Script Dependencies   #### -->

<!--  Packages -->  
```{r, echo=FALSE}
library(ggplot2)
library(plyr)
library(nlme)
library(lme4)
library(lmerTest)
```


<!--  Working Directory  -->       
```{r, echo=FALSE}
WD="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/LMMwREint"
setwd(WD)
```


<!-- Data  &  Variables -->       
```{r, echo=FALSE}
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/alldatFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/flowFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/indseqFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/mdataFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/seqFilter.RData")
```
***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
```{r, echo=FALSE}
index.malat1=which(rownames(seqFilter)=="MALAT1")
index.cd19=which(rownames(seqFilter)=="CD19")
mala=seqFilter[index.malat1,]
cd19=seqFilter[index.cd19,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=as.factor(mdataFilter$subject.no)
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd19, mala)
dat$logcd19=log(cd19+1, base = exp(1))
dat$logmala=log(mala+1, base = exp(1))
```

\vspace{10pt}

# Non-Transformed Model

```{r}
LMMwREint=lmer(mala~cd19+(1|subject.no), dat)
print(summary(LMMwREint))
print(AIC(LMMwREint))

coefInt=rep(fixef(LMMwREint)[1], times=15)
coefSlp=rep(fixef(LMMwREint)[2], times=15)

randInt=ranef(LMMwREint)
randInt=randInt$subject.no
randInt=randInt[,1]

coefInt=coefInt+randInt

plot.new()
plot(mala~cd19, data=dat, main="Model v Orignal Data")
for(i in 1:15)
{
  abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)
}
abline(a=fixef(LMMwREint)[1], b=fixef(LMMwREint)[2], lwd=4, col="red")

print("E[intercept|b_0=0]:")
print(fixef(LMMwREint)[1])
print("E[slope|b_0=0]:")
print(fixef(LMMwREint)[2])

dat$fit=fitted(LMMwREint)
dat$res=residuals(LMMwREint)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p

q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q
```


# Transformed Model
```{r}
logLMMwREint=lmer(logmala~logcd19+(1|subject.no), dat)

print(summary(logLMMwREint))
print(AIC(logLMMwREint))

coefInt=rep(fixef(logLMMwREint)[1], times=15)
coefSlp=rep(fixef(logLMMwREint)[2], times=15)

randInt=ranef(logLMMwREint)
randInt=randInt$subject.no
randInt=randInt[,1]

coefInt=coefInt+randInt

plot.new()
plot(logmala~logcd19, data=dat, main="Model v Orignal Data")
for(i in 1:15)
{
  abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)
}
abline(a=fixef(logLMMwREint)[1], b=fixef(logLMMwREint)[2], lwd=4, col="red")

print("E[intercept|b_0=0]:")
print(fixef(logLMMwREint)[1])
print("E[slope|b_0=0]:")
print(fixef(logLMMwREint)[2])

dat$fit=fitted(logLMMwREint)
dat$res=residuals(logLMMwREint)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p

q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q
```

<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

