---
title: "Linear Model with Fixed Effect Intercept"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce, and analyze a linear model with a fixed-effect subject-specific intercept, on the finalized data.  It will first produce a single response-predictor model for seq variables, and then analyze the assumptions made for OLS modeling.

***

<!--  ####  Script Dependencies   #### -->


<!--  Packages -->  
```{r, echo=FALSE}
library(ggplot2)
library(plyr)
library(nlme)
```


<!--  Working Directory  -->       

```{r, echo=FALSE}
WD="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/LMwFEint"
setwd(WD)
```


<!-- Data  &  Variables -->       
```{r, echo=FALSE}
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/alldatFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/flowFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/indseqFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/mdataFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/seqFilter.RData")
```
***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>

# Initial Models (Sequencing Variables)

## Response = mala, Predictor = cd19

```{r, echo=FALSE}
index.malat1=which(rownames(seqFilter)=="MALAT1")
index.cd19=which(rownames(seqFilter)=="CD19")
mala=seqFilter[index.malat1,]
cd19=seqFilter[index.cd19,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=mdataFilter$subject.no
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd19, mala)
dat$logcd19=log(cd19+1, base = exp(1))
dat$logmala=log(mala+1, base = exp(1))
```

\vspace{10pt}
# Model 0

## Non-Transformed Model
```{r, fig.width=4}
lmod0=lm(mala~cd19, dat)
(lmod0s=summary(lmod0))
plot(lmod0)
```

```{r}
p=ggplot(dat, aes(x=cd19, y=mala, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_abline(intercept = coefficients(lmod0)[1], slope = coefficients(lmod0)[2], color="BLUE")+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p
```


## Transformed Model
```{r, fig.width=4}
loglmod0=lm(logmala~logcd19, dat)
(loglmod0s=summary(loglmod0))
plot(loglmod0)
```

```{r}
p=ggplot(dat, aes(x=logcd19, y=logmala, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_abline(intercept = coefficients(loglmod0)[1], slope = coefficients(loglmod0)[2], color="BLUE")+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p
```


# Model 1
```{r}
LMwFEint=lm(mala~subject.no + cd19 , data = dat)
print(summary(LMwFEint))
print(AIC(LMwFEint))

coefIntercept=rep(coefficients(LMwFEint)[1], times=15)
coefSlope=rep(coefficients(LMwFEint)[16], times=15)

for( i in 2:15){
  coefIntercept[i]=coefIntercept[i]+coefficients(LMwFEint)[i]
}

plot(mala~cd19, data=dat, main="Model v Orignal Data")
for(i in 1:15)
{
  abline(a=coefIntercept[i], b=coefSlope[i], col=(20+i), lty=2, lwd=2)
}

dat$fitted=fitted(LMwFEint)
dat$resid=residuals(LMwFEint)

p=ggplot(dat, aes(x=fitted, y=resid, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p

q=ggplot(dat, aes(sample = resid))+
  stat_qq()+
  stat_qq_line()
q
```
# Log-Model 1
```{r}
logLMwFEint=lm(logmala~subject.no + logcd19 , data = dat)
print(summary(logLMwFEint))
print(AIC(logLMwFEint))

coefIntercept=rep(coefficients(logLMwFEint)[1], times=15)
coefSlope=rep(coefficients(logLMwFEint)[16], times=15)

for( i in 2:15){
  coefIntercept[i]=coefIntercept[i]+coefficients(logLMwFEint)[i]
}

plot(logmala~logcd19, data=dat, main="Model v Orignal Data")
for(i in 1:15)
{
  abline(a=coefIntercept[i], b=coefSlope[i], col=(20+i), lty=2, lwd=2)
}

dat$fitted=fitted(logLMwFEint)
dat$resid=residuals(logLMwFEint)

p=ggplot(dat, aes(x=fitted, y=resid, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p

q=ggplot(dat, aes(sample = resid))+
  stat_qq()+
  stat_qq_line()
q
```

<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

