---
title: "Linear Model with Fixed Effect Slope"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->
\begin{center}
\LARGE{\underline{\textbf{Description}}}
\end{center}

This script will produce, and analyze a linear model with a fixed-effect subject-specific intercept,  and a fixed-effect subject-specific slope, on the finalized data.  It will first produce a single response-predictor model for seq variables, and then analyze the assumptions made for OLS modeling.

***

<!--  ####  Script Dependencies   #### -->


<!--  Packages -->  
```{r, echo=FALSE}
library(ggplot2)
library(plyr)
library(nlme)
library(emmeans)
library(gridExtra)
```


<!--  Working Directory  -->       
```{r, echo=FALSE}
WD="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/LMeFEslope"
setwd(WD)
```


<!-- Data  &  Variables -->       
```{r, echo=FALSE}
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/alldatFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/flowFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/indseqFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/mdataFilter.RData")
load(file="/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/FinalData/InitialModels/InterceptOnly/seqFilter.RData")
```
***
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>
```{r, echo=FALSE}
index.malat1=which(rownames(seqFilter)=="MALAT1")
index.cd19=which(rownames(seqFilter)=="CD19")
mala=seqFilter[index.malat1,]
cd19=seqFilter[index.cd19,]
nFeature=mdataFilter$nFeature
nCount=mdataFilter$nCount
Perc.Mt=mdataFilter$Perc.Mt
subject.no=mdataFilter$subject.no
measurement=mdataFilter$measurement.name
dat=data.frame(subject.no, measurement, Perc.Mt, nCount, nFeature, cd19, mala)
dat$logcd19=log(cd19+1, base = exp(1))
dat$logmala=log(mala+1, base = exp(1))
```

\vspace{10pt}
# Non-Transformed Model
```{r}
LMwFEslope=lm(mala~subject.no + cd19:subject.no + cd19 , data = dat)
print(summary(LMwFEslope))
print(AIC(LMwFEslope))

coefInt=rep(coefficients(LMwFEslope)[1], times=15)
coefSlp=rep(coefficients(LMwFEslope)[16], times=15)

coef.int=coefficients(LMwFEslope)[2:15]
coef.int=c(0,coef.int)

coef.slopes=coefficients(LMwFEslope)[17:30]
coef.slopes=c(0,coef.slopes)

coefInt=coefInt+coef.int
coefSlp=coefSlp+coef.slopes

rgLMwFEslope=ref_grid(LMwFEslope)
lsmeans(rgLMwFEslope, "cd19")
contrast(rgLMwFEslope, interaction = T)


plot(mala~cd19, data=dat, main="Model v Orignal Data")
for(i in 1:15){
  abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)
}

dat$fit=fitted(LMwFEslope)
dat$res=residuals(LMwFEslope)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  ylab("Residual Value")
p


q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q

```

This fitted v residual plot "different" from the first model, but still does not exhibit any symmetrical variation about "0" (or any baseline) which is a VERY strong indicator that residuals are NOT normally distributed.  

# Transformed Model
```{r}
logLMwFEslope=lm(logmala~subject.no + subject.no:logcd19 + logcd19 , data = dat)

coefTable.lmod2=coef(logLMwFEslope)



print(summary(logLMwFEslope))
print(AIC(logLMwFEslope))

coefInt=rep(coefficients(logLMwFEslope)[1], times=15)
coefSlp=rep(coefficients(logLMwFEslope)[16], times=15)

coef.int=coefficients(logLMwFEslope)[2:15]
coef.int=c(0,coef.int)

coef.slopes=coefficients(logLMwFEslope)[17:30]
coef.slopes=c(0,coef.slopes)

coefInt=coefInt+coef.int
coefSlp=coefSlp+coef.slopes

plot(logmala~logcd19, data=dat, main="Model v Orignal Data")
for(i in 1:15){
  abline(a=coefInt[i], b=coefSlp[i], col=(20+i), lty=2, lwd=2)
}

dat$fit=fitted(logLMwFEslope)
dat$res=residuals(logLMwFEslope)

p=ggplot(dat, aes(x=fit, y=res, color=subject.no))+
  geom_point()+
  geom_hline(yintercept = 0)+
  ggtitle("Fitted Vs Residual")+
  xlab("Fitted Value")+
  theme(legend.position = "none")+
  ylab("Residual Value")
p


q=ggplot(dat, aes(sample = res))+
  stat_qq()+
  stat_qq_line()
q

grid.arrange(p,q,nrow=2)
```

<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

