---
title: "Model Summaries"
output: 
  pdf_document:
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=0.5in
fontsize: 11pt
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```
***

<!--  ####  Description   #### -->

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load("/Users/lee/Documents/Lee/School/CU Denver/MS_Project/Data:Scripts/RBC/Modeling/ModelSummaries/CompiledModelWorkspace.RData")
```
\begin{center}
\LARGE{\underline{NON-Tranformed Models}}
\end{center}



# Summary of Intercept and Slope Parameters

## Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-value  \\
\hline
\hline    
M0 (lmod0) & 7.8294e+03 & 2.3480e+02 & 33.344 &  <2e-16  \\
\hline
M1 (LMwFEint) & 10.1124e+03 & 8.4495e+02 & 11.968 & <2e-16 \\
\hline  
M2 (LMwFEslope) & 9.559e+03 & 8.956e+02 & 10.673 & <2e-16  \\
\hline
M3 (LMMwREint) & 7.3994e+03 & 7.6899e+02 & 9.622 & 8.06e-08 \\
\hline
M4 (LMMwREslope) & 7.4050e+03 & 7.6776e+02 & 9.645 & 8.4e-08 \\
\hline
M5 (GEEmod) & 7.8294e+03 & 9.8883e+02 & 62.692** & 2.44e-15** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)

## Main-Effect Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-Value  \\
\hline
\hline
M0 (lmod0) & 0.7132 & 1.5426  & 0.462 & <2e-16 \\
\hline
M1 (LMwFEint) & 2.562 & 1.501  & 1.707 & 0.0881 \\
\hline
M2 (LMwFEslope) & 1.765e+01 & 8.246  &  2.140 & 0.03256 \\
\hline
M3 (LMMwREint) & 2.495  & 1.491 & 1.674 & 0.0945  \\
\hline
M4 (LMMwREslope)  & 2.168 &  1.797  & 1.206 & 0.282 \\
\hline
M5 (GEEmod) & 0.7132 & 2.5028  & 0.081** & 0.776**  \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)



# Model Summaries

## M0 (lmod0)
```{r}
summary(lmod0)
```

## M1 (LMwFEint)
```{r}
summary(LMwFEint)
```

## M2 (LMwFEslope)
```{r}
summary(LMwFEslope)
```

## M3 (LMMwREint)
```{r}
summary(LMMwREint)
```

## M4 (LMMwREslope)
```{r}
summary(LMMwREslope)
```

## M5 (GEEmod)
```{r}
summary(GEEmod)
```

# Nested Model Comparisons
```{r}
# M0 < M1
anova(lmod0, LMwFEint )

# M1 < M2
anova(LMwFEint, LMwFEslope)

# M3 < M4
anova(LMMwREint, LMMwREslope)
```

# AIC Measurement Statistics
```{r}
AIC(lmod0,LMwFEint, LMwFEslope, LMMwREint, LMMwREslope)
```
***

\begin{center}
\LARGE{\underline{Tranformed Models}}
\end{center}

# Summary of Intercept and Slope Parameters

## Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-Value \\
\hline
\hline
M0 (loglmod0) & 8.46182 & 0.04568 & 185.258 & < 2e-16 \\
\hline
M1 (logLMwFEint) & 7.54861 & 0.12261 & 61.564 & < 2e-16 \\
\hline
M2 (logLMwFEslope) & 6.97932 & 0.13896 & 50.226 & < 2e-16 \\
\hline
M3 (logLMMwREint) & 8.41374 & 0.11825 & 71.151 & < 2e-16 \\
\hline
M4 (logLMMwREslope) & 8.39722 & 0.13957 & 60.166 & <2e-16 \\
\hline
M5 (logGEEmod) & 8.4618 & 0.1842 & 2109.82** & <2e-16** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)


## Main-Effect Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Model Number & Estimate & Std. Error & t-Stat & p-value \\
\hline
\hline
M0 (loglmod0) & 0.04918 & 0.01455  & 3.381 & 0.000747 \\
\hline
M1 (logLMwFEint) & 0.04833 & 0.01381  & 3.500 & 0.000484 \\
\hline
M2 (logLMwFEslope) & 0.51428 & 0.06017  &  8.546 & < 2e-16 \\
\hline
M3 (logLMMwREint) & 0.04920  & 0.01374 & 3.579  & 0.00036 \\
\hline
M4 (logLMMwREslope)  & 0.05938 &  0.03538  & 1.678 & 0.119 \\
\hline
M5 (logGEEmod) & 0.0492 & 0.0397  & 1.53** & 0.22** \\
\hline
\end{tabular}
\end{center}
Note: ** These are Wald test of a single parameter (not t-tests)



# Model Summaries

## M0 (loglmod0)
```{r}
summary(loglmod0)
```

## M1 (logLMwFEint)
```{r}
summary(logLMwFEint)
```

## M2 (logLMwFEslope)
```{r}
summary(logLMwFEslope)
```

## M3 (logLMMwREint)
```{r}
summary(logLMMwREint)
```

## M4 (logLMMwREslope)
```{r}
summary(logLMMwREslope)
```

## M5 (logGEEmod)
```{r}
summary(logGEEmod)
```

# Nested Model Comparisons
```{r}
# M0 < M1
anova(loglmod0, logLMwFEint )

# M1 < M2
anova(logLMwFEint, logLMwFEslope)

# M3 < M4
anova(logLMMwREint, logLMMwREslope)
```


# AIC Measurement Statistics
```{r}
AIC(loglmod0, logLMwFEint, logLMwFEslope, 
    logLMMwREint, logLMMwREslope)
```














<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Begin Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- End Script -->       
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
<!-- Post-Script -->       
<!------------------------------------------------------------------------------>

<!-- Notes:        -->       

<!-- Compilation Errors -->       

<!-- Execution Errors -->

<!-- Next Scripts to Consider -->

<!------------------------------------------------------------------------------>

