---
title: "Comparing Models of Subject-Clustered Single-Cell Data"
author: "Lee Panter"
csl: mathematical-biosciences-and-engineering.csl
output:
  pdf_document: 
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
  word_document: 
geometry: margin=1.0in
fontsize: 12pt
subtitle: Version 6.0-Results
bibliography: Bib_AccRefSheet.bib
---


<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/v6/5 Results"
setwd(WD)
```
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
# Results

Table 8 and table 9 display parameter value estimates, standard errors, test statistics, and p-values for the main-effect slope term estimated by all five modeling approaches:

## Coefficient Estimates

### (MALAT1 ~ CD19)

\small{
\begin{center}
\centering
\begin{tabular}{|m{0.15\textwidth}|m{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  4.918e-2 &  1.455e-2  & 3.381* &  7.47e-4\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  4.833e-2 &  1.381e-2 & 3.500* &  4.84e-4\\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  4.920e-2 &  1.374e-2 & 3.579*  &  3.6e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 5.938e-2 &  3.538e-2  & 1.678* &  1.19e-1 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 4.918e-2 & 1.455e-2 & 3.381** & 7.47e-4 \\
\hline
\end{tabular}

\vspace{5pt}

Table 8: Summary Table for $CD19 \sim MALAT1$ variable parings.  * Approximate normal distribution.   ** Approximate Wald-Z distribution
\end{center}
}


### (FBLN1 ~ CD34)
\small{
\begin{center}
\centering
\begin{tabular}{|m{0.15\textwidth}|m{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  7.884e-1 &  4.92e-2  & 4.002* & <2e-16\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  1.31e-1 & 3.42e-2  & 3.818* & 1.42e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  1.35e-1 & 3.42e-2  & 3.95* & 8.4e-5 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 1.705e-1  & 7.29e-2 & 2.34*  & 6.7e-2 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 7.884e-1 & 4.92e-2 & 4.002** & < 2e-16 \\
\hline
\end{tabular}

\vspace{5pt}

Table 9: Summary Table for $CD34 \sim FBLN$ variable parings.  * Approximate normal distribution.   ** Approximate Wald-Z distribution
\end{center}

}

The main-effect slope parameter represents subject-inspecific information about how predictor and response variables are correlated. We have seen that each method accommodates the effects of subject-level correlation differently. Specifically, we noted that the Linear Model and GEE methods estimated population-averaged parameters, whereas the other models had subject-specific interpretations. So, when the main-effect slope parameter estimate is compared across methods with otherwise identical in structure, we can directly attribute changes to a shift in this parameters value to be a redistribution of the attributed source of correlation between the variables of interest.

The percent change in the main-effect slope parameter across models is displayed for each variable paring in Tables 10 and 11. Values are full-percentage changes, and are calculated using:

$$\text{Percent Change} [A]_{ij} = \left(\frac{A_{j}-A_{i}}{A_{i}}\right) * 100$$

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0 &  -1.7283 &  0.0407 & 20.7401 &  0.0000 \\
\hline
LM-FE & 1.7587 &  0 &  1.8001 & 22.8636  & 1.7587  \\      
\hline
LMM-RI & -0.0407 & -1.7683 &  0 & 20.6911 & -0.0407  \\
\hline
LMM-RS & -17.1775 & -18.6090 & -17.1438 & 0 & -17.1775 \\
\hline
GEE & 0.0000 & -1.7283  & 0.0407 & 20.7401  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table 10: Main effect slope Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}

\vspace{10pt}

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -83.3841 & -82.8767  & -78.3739 & 0.0000  \\
\hline
LM-FE & 501.8321 & 0   & 3.0534    & 30.1527  & 501.8321 \\
\hline
LM-RI & 484.0000 & -2.9630  & 0    & 26.2963  & 484.0000  \\
\hline
LM-RS & 362.4047 & -23.1672 & -20.8211  & 0   & 362.4047  \\
\hline
GEE & 0.0000   & -83.3841 & -82.8767  & -78.3739 & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 11: Main effect slope Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}

It is worthwhile to comment on the consistency properties of estimates across models within variable parings. In each of the variable paring scenarios we see that changes between models within either of the cases:

1. $\text{LM} \Leftrightarrow \text{GEE}$ (identical estimates/zero percent change)
2. $\text{LM-FE} \Leftrightarrow \text{LMM-RI}\Leftrightarrow \text{LMM-RS}$

results in smaller percent-change values than changes between the cases. Changes between LMM-RS and LM-FE/LMM-RI in the CD19-MALAT1 variable paring are technically higher, but this result is most likely an artifact of subject-specific interactions with the covariate. Since models within each of the cases (1) and (2) above estimate similarly interpreted parameters (subject-specific vs population averaged) the result is otherwise expected.

## Standard Error Estimates

The standard errors for this parameter are also enlightening when compared across models. A change in a parameter estimate’s standard error across modeling methodology represents a revision in the underlying distributional conclusions the method is using to support its result. In this way, an increased standard error between two models that are estimating the same parameter indicates a decrease in obtained (obtainable) estimate precision.

Tables 12 and 13 are percent change matrices for the standard error of the main effect slope parameter:


\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -5.0859  & -5.5670   & 143.1615  & 0.0000  \\
\hline
LM-FE & 5.3584   & 0   & -0.5069   & 156.1912  & 5.3584  \\     
\hline
LMM-RI & 5.8952   & 0.5095   & 0    & 157.4964  & 5.8952  \\
\hline
LMM-RS & -58.8751 & -60.9666 & -61.1645  & 0    & -58.8751  \\
\hline
GEE &  0.0000   & -5.0859  & -5.5670   & 143.1615  & 0        \\
\hline
\end{tabular}

\vspace{5pt}

Table 12: Main effect slope Standar Error Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}


\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -30.4878 & -30.4878  & 48.1707   & 0.0000  \\
\hline
LM-FE & 43.8596  & 0   & 0.0000    & 113.1579  & 43.8596  \\
\hline
LM-RI & 43.8596  & 0.0000   & 0    & 113.1579  & 43.8596  \\
\hline
LM-RS & -32.5103 & -53.0864 & -53.0864  & 0    & -32.5103  \\
\hline
GEE &  0.0000   & -30.4878 & -30.4878  & 48.1707   & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 13: Main effect slope Standar Error Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}


Changes in standard errors display similarly infomative consistencies.  In each variable pairing:

1. The standard error increases on the following model transitions:
    a. $\text{All Other Models } \longrightarrow \text{LMM-RS}$
    b. $\text{LMM-RI} \longrightarrow \text{All Other  Models}$



2. The standard error decreases on the following model transitions:
    a. $\text{All Other Models} \longrightarrow \text{LMM-RI}$
    b. $\text{LMM-RS} \longrightarrow \text{All other Models}$
    c. $\text{LM} \longrightarrow \text{LM-FE}$
    d. $\text{GEE} \longrightarrow \text{LM-FE}$




a. The modeling transitions in (1a) correspond with the addition of information to the model in the form of a subject-specific "Random Effect Slope". 
b. The transitions in (1b) correspond to either:
    i. addition of the parameter in 1a
    ii. loss of subject-specific information that was originally incorperated into the variance-component of the model.  I.e., loss of subject-specific variability information.
c. The transitions in (2a and 2b) are the inverse representation of the relationships outlined in (a) and (b) above.
d. The transition in (2c) corresponds to the incorperation of additive, subject-specific, predictor independent (mean-effect) information into the model.
f. The transition in (2d) corresponds to the addition of the assumption of independence between subjects, along with a purely parametric fitting method (for LM-FE). 

The preceding relationships allow us to deduce the effects of the various types of information inclusion on the precision of parameters used to make inferences on the relationship between predictor and response. Beneficial (increases in precision) information inclusions will result in reductions to standard error estimates (section 2 transitions, with explanations c and d above). Detrimental (decreases in precision), or contradictory information will result in increased standard error estimates (section 1 transitions, a & b explanations).

Explanation (e) demonstrates the importance of considering the effect of correlation between subjects in single-cell data. Since this transition is non-zero, it is clear that there is an effect associated with subject-clustered sampling. Otherwise the use of an independence assumption in an analysis would lead to identical results as an analysis without this assumption.


## Test Statistics

Tables 14 and 15 are percent change matrices for the test statistic of the main effect slope parameter:

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & 3.5197   & 5.8563    & -50.3697  & 0.0000 \\
\hline
LM-FE & -3.4000  & 0   & 2.2571    & -52.0571  & -3.4000 \\
\hline
LMM-RI & -5.5323  & -2.2073  & 0    & -53.1154  & -5.5323  \\
\hline
LM-RS & 101.4899 & 108.5816 & 113.2896  & 0    & 101.4899 \\
\hline
GEE & 0.0000  & 3.5197   & 5.8563    & -50.3697  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table 14: Main effect slope Test Statistic Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}



\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & -4.5977  & -1.2994   & -41.5292  & 0.0000  \\
\hline
LM-FE & 4.8193  & 0   & 3.4573    & -38.7114  & 4.8193  \\
\hline
LM-RI & 1.3165  & -3.3418  & 0    & -40.7595  & 1.3165  \\
\hline
LM-RS & 71.0256 & 63.1624  & 68.8034   & 0    & 71.0256  \\
\hline
GEE & 0.0000  & -4.5977  & -1.2994   & -41.5292  & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 15: Main effect slope Test Statistic Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}

If we look back to Tables 8 and 9, we can see that the sign of the test statistic for each model remains positive across each method, as well as across each variable pairing. The test-statistic distributions, which are approximately normal, or Wald-Z (asymptotically z-distributed) can be used to justify a symmetry argument that larger-valued test statistics represent higher significance of the parameter. This is analogous to higher magnitude, normally distributed, test statistics representing higher significance.

The patterns that we observe in the test statistic percent change matrices serves to largely reinforce previous conclusions we have made using the estimates of coefficients or standard errors. Transitions between LM/GEE and LM/LMM-RI tend to result in larger test statistic changes than changes within models that estimate similar parameters (i.e LM-FE $\leftrightarrow$ LMM-RI which both estimate subject-specific parameters, and LM $\leftrightarrow$ GEE which both estimate population-averaged parameters).

Transitions to LMM-RS from any model results in a loss of significance of the main effect slope parameter as indicated by negative test statistic percent changes. This aligns with our intuition as we would expect the average relationship between outcome and covariate to diminish as emphasis on subject-specific relationship between outcome and covariate increased.

The results outlined in the section above are all based on the inclusion of various types of subject-specific information. These relationships can be classified according to how they affect our ability to perform inference on the relationship between a predictor and a response using subject-correlated scRNA-seq data. To this effect, we can now evaluate our variable-pairing relationship(s) to determine if there is a significant effect from the nested sampling methods used to create the scRNA-seq data, and if there is an effect, how can this effect best be accounted for.

<!------------------------------------------------------------------------------>


# References

\bibliography{Bib_AccRefSheet}

