---
csl: mathematical-biosciences-and-engineering.csl
output:
  pdf_document: 
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
  word_document: 
geometry: margin=1.0in
fontsize: 12pt
subtitle: Version v2.2-Model Descriptions
bibliography: Bib_AccRefSheet.bib
---


<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/V2.2/4 Model Descriptions"
setwd(WD)
```
<!------------------------------------------------------------------------------>

<!------------------------------------------------------------------------------>

## Generalized Estimating Equations (GEE)

The final modeling method considered is generalized estimating equations (GEE). The GEE method is an extension of the methods used to estimate Generalized Linear Models (GLMs) to categorical type responses [@Introduction2013GeneralizedEstimatingEquations].  

focuses on estimating the fixed effect parameters specified within a model.  The method does not directly estimate variances and covariances, but instead approximates these values using a "working variance-covariance" structure. 

Procedurally GEE estimates are computed by finding numerical solutions for
an \textit{estimating equation} $U(\mathbf{\beta})=0$, where

\begin{equation}
U(\mathbf{\beta})=\sum_{i=1}^{N} \left \{\mathbf{D}_{i}^{T}\mathbf{V}_{i}^{-1}\left(\mathbf{y}_{i} - \mathbf{\mu}_{i}(\beta) \right)   \right \}
\end{equation}

and
\begin{align*}
g \left(\mathbf{\mu}_{i} (\beta) \right) &=\mathbf{\eta}_{i} = E\left [\mathbf{Y}_{i} | X_{i} \right ]\\
&= X_{i}^{T} \mathbf{\beta}
\end{align*} 
outlines the dependence of the marginal expectation of the response on the covariates through the link function $g()$, and therefore:
\begin{align*}
\mathbf{\mu}_{i} (\beta)  &= g^{-1} \left(\mathbf{\eta}_{i}\right)  = g^{-1} \left(E\left [\mathbf{Y}_{i} | X_{i} \right ] \right) \\
&= g^{-1} \left(X_{i}^{T} \mathbf{\beta} \right) 
\end{align*}
The first derivative matrix $\mathbf{D_{i}}$ is given by:
$$\mathbf{D}_{i}=
\begin{bmatrix}
\frac{\partial \mu_{i1}}{\beta_{1}} & \frac{\partial \mu_{i1}}{\beta_{2}} & \cdots & \frac{\partial \mu_{i1}}{\beta_{p}}\\
\frac{\partial \mu_{i2}}{\beta_{1}} & \frac{\partial \mu_{i2}}{\beta_{2}} & \cdots & \frac{\partial \mu_{i2}}{\beta_{p}}\\
\vdots & \vdots & \ddots & \vdots \\
\frac{\partial \mu_{in_{i}}}{\beta_{1}} & \frac{\partial \mu_{in_{i}}}{\beta_{2}} & \cdots & \frac{\partial \mu_{in_{i}}}{\beta_{p}}\\
\end{bmatrix}$$

$\mathbf{V}_{i}$ is the working covariance matrix.  This matrix can be estimated from hyper parameters created in the process of numerically solving for the GEE estimates $\beta_{GEE}$ that solve $U\left(\beta\right)=0$. Please see Fitzmaurice, Laird, and Ware [@fitzmaurice2012applied] for more information.

The GEE algorithm uses the following general steps to converge at an estimate:

1. Generalized linear modeling methods employing maximum likelihood estimation are used to obtain intial estimates for $\mathbf{\beta}$
2. Estimates for $\mathbf{\beta}$ (from 1) used to compute hyper-parameters
3. New estimates for hyper-parameters and working covariance matrix ($\mathbf{V}_{i}$) used to obtain new estimates for $\mathbf{\beta}$ by solving (1)
4. Repeat Steps 2 & 3 until algorithm converges

The algorithm is robust to misspecification of the observational covariance structure.  So initially incorrect specifications of the working covariance matrix still converge to the appropriate structure form with algorithmic iteration.   

The GEE algorithm is stable, in-part due to the fact that the method estimates population-average effects.  Each of the previous methods (model LM withstanding) have subject-specific interpretations, but the GEE algorithm provides marginal parameter estimates.  These values do not represent any specific subject, but rather a \textit{hypothetical} average subject.

According to Fitzmaurice, Laird, and Ware [@fitzmaurice2012applied], responses modeled with the GEE process need to be stationary, i.e:

$$E\left [ Y_{ij} | \mathbf{X}_{i} \right ]=E\left [ Y_{ij} | X_{i1}, \ldots, X_{in_{i}}\right ]=E\left [ Y_{ij} | X_{ij}\right ]$$
The scRNA-seq data is assumed to be independent within-subject, therefore:

$$E\left [Y_{ij} | X_{ij}\right ] = E\left [Y_{ij} | X_{ij'}\right ]$$
$$\forall j \in \left \{  1, \ldots, n_{i}\right \}  \quad j\neq j'$$
as needed.

The three-part specification of the GEE framework includes:

1. The link function and linear predictor
2. Variance function
3. A working covariance matrix 

All three items are chosen so that the resulting model estimates are comparable to preceding estimates for intercept and slope. Therefore, the identity link function will be assumed:
$$g(x)=x$$
in conjunction with the linear predictor:
$$g(\mu_{i})=\eta_{i} = \beta_{0} + \beta_{1}X_{i}$$
which implies the modeling equation:
$$E\left [Y_{ij} \right ]=\mu_{ij}=\eta_{ij}= \beta_{0} + \beta_{1}X_{ij}$$
Additionally, an identity variance function will be assumed:
\begin{align*}
Var\left(Y_{ij} \right) &= \phi \nu\left(\mu_{ij}   \right)\\
&= \phi \mu_{ij}
\end{align*}
and a working covariance matrix structure for repeated measures that correspond to the assumption of independence of observations within a subject will be specified as:
$$\left [ Cov\left(Y_{ij}, Y_{ik}\right)\right ]_{jk}=
\begin{cases}
Var\left(Y_{ij}\right)  \quad &\mbox{if} \quad j=k \\
0 \quad &\mbox{if} \quad j \neq k \\
\end{cases}$$
$$for \quad j,k \in \left \{  1, \ldots, n_{i}   \right \}$$


## Parameter Interpretations 

The GEE and LM modeling techniques are methods of obtaining estimates of population-averaged parameters.  These parameter values are interpreted as contributing to the response of the \textit{hypothetical} average subject  (not representative of any single subject within the sample).

Suppose that $\hat{\beta}_{population}$ represents an estimate obtained for the fixed effect slope as obtained by one of the previously describe \textit{population-averaged} modeling methods.  An interpretation of this parameter is:  **across all subjects, a one-unit increase in the predictor ($X_{ij}$) is associated with a $(\hat{\beta}_{population})$ unit change in the expected outcome ($Y_{ij}$)**

Conversely, the LM-FE, LMM-RI and LMM-RS modeling techniques are methods of obtaining estimates of subject-specific parameters.  These parameter values are interpreted as contributing to the average response having controlled for a constant subject of origin (i.e. the parameter estimate attributable to a single subject within the sample) 
  
Suppose that $\hat{\beta}_{subject}$ represents an estimate obtained for the fixed effect slope as obtained by one of the previously describe \textit{subject-specific} modeling methods.  An interpretation of this parameter is: **for a given subject (controlling for the subject of origin) a one-unit increase in the predictor ($X_{ij}$) is associated with a $(\hat{\beta}_{subject})$ unit change in the expected outcome ($Y_{ij}$).**



<!------------------------------------------------------------------------------>


# References

\bibliography{Bib_AccRefSheet}



