---
csl: mathematical-biosciences-and-engineering.csl
output:
  word_document: 
  pdf_document: 
    df_print: kable
    includes:
      in_header: Rmarkdown_preamble.tex
geometry: margin=1.0in
fontsize: 12pt
subtitle: Version 1.2-Results
bibliography: Bib_AccRefSheet.bib
---


<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/V1.2/5 Results"
setwd(WD)
```
<!------------------------------------------------------------------------------>


<!------------------------------------------------------------------------------>
# Results

- The table below (Table _) shows the estimate, standard error, test statistic, and p-values associated with the fixed-effect slope parameter as estimated by each modeling method.

\small{
\begin{center}
\centering
\begin{tabular}{|m{0.15\textwidth}|m{0.15\textwidth}|m{0.185\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & \center{Variable Pairing} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline

\multirow{2}{*}{LM} & \multirow{2}{*}{Linear Model} & MALAT1-CD19  & 4.918e-2 &  1.455e-2  & 3.381 &  7.47e-4 \\
& & FBLN1-CD34   &  7.884e-1 &  4.92e-2  & 4.002 & <2e-16\\
\hline

\multirow{2}{*}{LM-FE} & \multirow{2}{*}{Linear Model} & MALAT1-CD19  &  4.833e-2 &  1.381e-2 & 3.500 &  4.84e-4\\
& Fixed Effect \newline Intercept & FBLN1-CD34   &  1.31e-1 & 3.42e-2  & 3.818 & 1.42e-4 \\
\hline

\multirow{2}{*}{LM-RI} & \multirow{2}{*}{Linear Model} & MALAT1-CD19  &  4.920e-2 &  1.374e-2 & 3.579  &  3.6e-4 \\
& Random \newline Intercept & FBLN1-CD34   &  1.35e-1 & 3.42e-2  & 3.95 & 8.4e-5 \\
\hline

\multirow{2}{*}{LM-RS} & \multirow{2}{*}{Linear Model} & MALAT1-CD19  & 5.938e-2 &  3.538e-2  & 1.678 &  1.19e-1 \\
& Random Slope & FBLN1-CD34   & 1.705e-1  & 7.29e-2 & 2.34  & 6.7e-2 \\
\hline

\multirow{2}{*}{GEE} & \multirow{2}{*}{Generalized} & MALAT1-CD19  & 4.918e-2 & 1.455e-2 & 3.381* & 7.47e-4 \\
& Estimating Equations & FBLN1-CD34   & 7.884e-1 & 4.92e-2 & 4.002* & < 2e-16 \\
\hline

\end{tabular}
\vspace{5pt}

Table : Summary table for each model method, and each variable pairing. ** Approximate Wald-Z distribution
\end{center}
}

- The fixed effect slope parameter is being considered for comparison because:
  - It has been estimated in all models, so the effects on the parameter itself can be interpreted as effects due to model-approach changes
  - It will allow us to diagnose effects due to subject-level clustering when we compare properties of the parameter across different modeling techniques

- Displayed:
  - percent change in: parameter estimate, standard error, and test statistic for the MALAT1~CD19 variable pairing as defined by:

$$\text{Percent Change} [A]_{ij} = \left(\frac{A_{j}-A_{i}}{A_{i}}\right) * 100$$

\begin{center}
\begin{tabular}{|cc||c|c|c|c|c|}
\hline
Model & Measure & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
 & Estimate & 0 &  -1.7283 &  0.0407 & 20.7401 &  0.0000 \\
LM & SE & 0   & -5.0859  & -5.5670   & 143.1615  & 0.0000  \\
 & Test-Stat & 0  & 3.5197   & 5.8563    & -50.3697  & 0.0000 \\
\hline
 & Estimate & 1.7587 &  0 &  1.8001 & 22.8636  & 1.7587  \\     
LM-FE & SE & 5.3584   & 0   & -0.5069   & 156.1912  & 5.3584  \\   
 & Test-Stat & -3.4000  & 0   & 2.2571    & -52.0571  & -3.4000 \\
\hline
 & Estimate & -0.0407 & -1.7683 &  0 & 20.6911 & -0.0407  \\
LMM-RI & SE & 5.8952   & 0.5095   & 0    & 157.4964  & 5.8952  \\
 & Test-Stat & -5.5323  & -2.2073  & 0    & -53.1154  & -5.5323  \\
\hline
 & Estimate & -17.1775 & -18.6090 & -17.1438 & 0 & -17.1775 \\
LMM-RS & SE & -58.8751 & -60.9666 & -61.1645  & 0    & -58.8751  \\
 & Test-Stat & 101.4899 & 108.5816 & 113.2896  & 0    & 101.4899 \\
\hline
 & Estimate & 0.0000 & -1.7283  & 0.0407 & 20.7401  & 0 \\
GEE & SE & 0.0000   & -5.0859  & -5.5670   & 143.1615  & 0 \\
 & Test-Stat & 0.0000  & 3.5197   & 5.8563    & -50.3697  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table: percent change in parameter estimate, standard error, and test statistic for the MALAT1-CD19 variable

\end{center}

\vspace{10pt}

- Displayed:
  - percent change in: parameter estimate, standard error, and test statistic for the FBLN1~CD34 variable pairing 

\begin{center}
\begin{tabular}{|cc||c|c|c|c|c|}
\hline
Model & Measure & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
 & Estimate & 0   & -83.3841 & -82.8767  & -78.3739 & 0.0000  \\
LM & SE & 0   & -30.4878 & -30.4878  & 48.1707   & 0.0000  \\
 & Test-Stat & 0  & -4.5977  & -1.2994   & -41.5292  & 0.0000  \\
\hline
 & Estimate & 501.8321 & 0   & 3.0534    & 30.1527  & 501.8321 \\     
LM-FE & SE & 43.8596  & 0   & 0.0000    & 113.1579  & 43.8596  \\   
 & Test-Stat & 4.8193  & 0   & 3.4573    & -38.7114  & 4.8193  \\
\hline
 & Estimate & 484.0000 & -2.9630  & 0    & 26.2963  & 484.0000  \\
LMM-RI & SE &  43.8596  & 0.0000   & 0    & 113.1579  & 43.8596  \\
 & Test-Stat & 1.3165  & -3.3418  & 0    & -40.7595  & 1.3165  \\
\hline
 & Estimate & 362.4047 & -23.1672 & -20.8211  & 0   & 362.4047  \\
LMM-RS & SE & -32.5103 & -53.0864 & -53.0864  & 0    & -32.5103  \\
 & Test-Stat & 71.0256 & 63.1624  & 68.8034   & 0    & 71.0256  \\
\hline
 & Estimate & 0.0000   & -83.3841 & -82.8767  & -78.3739 & 0  \\
GEE & SE &  0.0000   & -30.4878 & -30.4878  & 48.1707   & 0  \\
 & Test-Stat & 0.0000  & -4.5977  & -1.2994   & -41.5292  & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table: percent change in parameter estimate, standard error, and test statistic for the FBLN1-CD34 variable

\end{center}


- Camparison of parameter estimates
  - We compare changes in the parameter estimate values because:
    - magnitude of change when comparing two model estimates can be indicative of the amount of subject-level correlation in the data
  - Findings:
    - Models LM and GEE have similar estimates
      -indicates that models with population-average interpretation parameters are more similar to each other
    - Models LM-FE, LM-RI, and LM-RS are also all similar 
    -indicates that models with subject-specific interpretation parameters are more similar to each other
    - Estimates are less similar when comparing between the two clusters of similarity mentionioned
    - Accounting for subject-specific outcomes, results in a different estimate than when modeling population-averages, i.e. there is an effect from subject-level correlation
    - The effect is more pronounced in the FBLN1~CD34 pairing than the MALAT1~CD19 pairing



## Standard Error Estimates

- Comparison of standard error of estimates
  - We compare changes in the standard error of estimates because:
    - a change in a parameter's SE across modeling methodologies represents a revision in the underlying evidence the method is using to support its result.
    - e.g. increase SE between two models indicates a decrease in an estimate's precision
  - Findings:
    - The standard error of the fixed effect slope parameter estimate is highest for the random slope model, which differs from all models with the incorperation of subject-specific variablility relationships between predictor and response
    - The SE is lowest for the random intercept model, which differs from all models either by lacking the incorperation of subject-specific varibility relationships between predictor and response (i.e. nested within LMM-RS model), or incorperating subject-specific variability information independent of predictors
    - The fixed-effect subject-specific intercept had a lower SE than either of the population-average parameter interpretaion models LM and GEE.  The LM-FE model incorperates subject-specific predictor-independent information into the mean-effect of the model.
    





## Test Statistics

- Comparison of test statistics
  - we compare changes in the test-statistics of estimates because:
    - patterns we observe in test statistic percent change matrices can serve to reinforce previous conclusions we have made using estimated coefficients and standard errors.
  - Findings:
    - Changes in test statistics for the fixed-effect main-effect slope were smaller between LM and GEE, as well as between LM-FE, LMM-RI, and LMM-RS with few exceptions.
    - There was undetectable change between LM and GEE
    - Changes to the LMM-RS resulted in loss of significance (larger p-value)
    
    
<!------------------------------------------------------------------------------>

## Nested Model Comparisons

### Test for FE Subject Specific Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & Resid DF & RSS & DF & Sum of Squares & F-stat & P(>F) \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LM & 1108 & 1167.76 & & & & \\
 & LM-FE & 1094 & 935.89 & 14 & 231.87 & 19.36 & < 2.2e-16 \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LM & 1108 & 650.51 & & & & \\
 & LM-FE & 1094 & 214.92 & 14 & 435.59 & 158.38 & < 2.2e-16 \\
\hline
\hline
\end{tabular}

\vspace{5pt}

Table \#: ANOVA nested model comparison table for testing the inclusion of the subject-specific fixed-effect intercept
\end{center}


The table above is a nested model comparison, the result of which is an F-test statistic telling us that there is very strong evidence to support the inclusion of the subject-specific fixed-effect intercept into the LM model. 




### Test for RE Subject Specific Intercept

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & df & AIC & logLik & L.Ratio & p-value \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LM & 3 & 3224.097 & -1609.048 &  & \\
 & LMM-RI & 4 & 3032.024 & -1512.012 & 194.0722 & <0.0001  \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LM & 3 & 2572.807 & -1283.403 &  & \\
 & LMM-RI & 4 & 1438.086 & -715.043 & 1136.72 & <0.0001  \\
\hline
\hline
\end{tabular}

\vspace{5pt}

Table \#: ANOVA nested model comparison table for testing the inclusion of the subject-specific random effect intercept
\end{center}


The table above is a nested model comparison, the result of which is a likelihood ratio statistic telling us that there is very strong evidence to support the inclusion of the subject-specific random effect intercept into the LM model. 



### Test for RE Subject Specific Slope

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Variable Pair & Model & df & AIC & logLik & L.Ratio & p-value \\
\hline
\hline
\multirow{2}{*}{MALAT1-CD19} & LMM-RI & 4 & 3032.024 & -1512.012 &  & \\
 & LMM-RS & 6 & 2993.820 & -1490.910 & 42.20503 & <0.0001  \\
\hline
\hline
\multirow{2}{*}{FBLN1-CD34} & LMM-RI & 4 & 1438.086 & -715.043 &  & \\
 & LMM-RS & 6 & 1438.068 & -713.034 & 4.018095 & 0.1341  \\
\hline
\end{tabular}

\vspace{5pt}

Table \#: ANOVA nested model comparison table for testing the inclusion of the subject-specific random effect slope
\end{center}


The table above is a nested model comparison, the result of which is a likelihood ratio statistic telling us that there is very strong evidence to support the inclusion of the subject-specific random effect slope into the LMM-RI model for the MALAT1~CD19 variable pairing. However, there is insufficient evidence to support the inclusion of the subject-specific random effect slope into the LMM-RI model for the FBLN~CD34 variable pairing.





# References

\bibliography{Bib_AccRefSheet}

