---
title: "Table Doc"
subtitle: "V51"
author: "Lee Panter"
output: pdf_document
---

<!------------------------------------------------------------------------------>
<!--  ####  KNITR Setup & Script Information   #### -->
<!------------------------------------------------------------------------------>

<!--  ####  KNITR Specs   #### -->
```{r setup, cache=TRUE, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo=FALSE, 
                      cache = TRUE, 
                      fig.align = "center",
                      fig.width = 5)
```


<!-- ####	 Set Working Directory	 ####   -->
```{r, echo=FALSE}
WD="/Users/lee/Documents/GitHub/MSproject_RBC/MSproject_RBC/FinalizedWriteUps/v5/5.1"
setwd(WD)
```
<!------------------------------------------------------------------------------>


\begin{figure}[!h]
\centering
		\includegraphics[width=0.75\textwidth]{2020-01-05_21-32-42.jpeg}
  \caption{Pre-QC $PerctMT$ Distribution for each subject}
\end{figure}

--------------------------------------------------------------------------

\newpage

<!------------------------------------------------------------------------------>

\begin{figure}[!h]
\centering
		\includegraphics[width=0.75\textwidth]{2020-01-05_21-32-14.jpeg}
  \caption{Post QC distribution of $PerctMT$ with thresholds implemented by Arazi, Rao, Berthier, et al}
\end{figure}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{figure}[!h]
\centering
		\includegraphics[width=0.85\textwidth]{2020-01-05_21-42-05.jpeg}
		\caption{Post QC distribution of $PerctMT$ with thresholds implemented in this paper}
\end{figure}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{flushleft}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|}
\hline
Subject Number & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
\hline
\hline
Number of Observations Before QC & 375 & 375 & 364 & 381 & 340 & 383 & 383 & 356 & 372 \\
\hline
Number of Observations After QC & 0 & 0 & 0 & 0 & 58 & 86 & 32 & 0 & 31 \\
\hline
\end{tabular}



\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
Subject Number & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 \\
\hline
\hline
Number of Observations Before QC & 327 & 311 & 379 & 375 & 345 & 371 & 381 & 381 & 377 & 380 \\
\hline
Number of Observations After QC & 21 & 107 & 0 & 107 & 100 & 25 & 0 & 122 & 0 & 127 \\
\hline
\end{tabular}



\begin{tabular}{|c||c|c|c|c|c|c|c|c|}
\hline
Subject Number & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 \\
\hline
\hline
Number of Observations Before QC & 381 & 380 & 333 & 333 & 239 & 218 & 378 & 342 \\
\hline
Number of Observations After QC& 75 & 0 & 87 & 0 & 79 & 0 & 53 & 0 \\
\hline
\end{tabular}
\end{flushleft}
\begin{center}
Table 1: Observation counts per-subject before and after Quality Control threshold filter restrictions
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{flushleft}
\begin{tabular}{|c||c|c|c|c|c|c|c|c|}
\hline
Subject Group Number & 5 & 6 & 7 & 9 & 10 & 11  & 13 & 14 \\
\hline
Number of Observations & 58 & 86 & 32 & 31 & 21 & 107  & 107 & 100\\
\hline
\end{tabular}

\vspace{5pt}


\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
Subject Group Number & 15  & 17  & 19 & 20  & 22  & 24  & 26 \\
\hline
Number of Observations & 25  & 122  & 127 & 75 & 87  & 79  & 53  \\
\hline
\end{tabular}

\vspace{5pt}

\end{flushleft}

\begin{center}
Table 2: Observation count per-subject, subjects with positive counts
\end{center}

\vspace{5pt}

Table 3 displays the descriptive statistics for the number of observations per-subject.

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
MIN & 1st Q & Median & Mean & 3rd Q & MAX \\
\hline
21 & 42.5 & 79 & 74.0 & 103.5 & 127 \\
\hline
\end{tabular}

\vspace{5pt}

Table 3: observation count per-subject descriptive statistics
\end{center}

<!------------------------------------------------------------------------------>


\newpage

<!------------------------------------------------------------------------------>

\begin{figure}[h!]
\centering
		\includegraphics[width=0.85\textwidth]{Untranslated.jpg}
		\caption{Predictor-Response pairing variable distributions}
\end{figure}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{figure}[h!]
\centering
		\includegraphics[width=0.85\textwidth]{Transformed.jpg}
		\caption{Predictor-Response variable pairings, post-transformation distributions}
\end{figure}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>


\begin{center}
\centering
\begin{tabular}{|p{0.15\textwidth}|p{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  4.918e-2 &  1.455e-2  & 3.381* &  7.47e-4\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  4.833e-2 &  1.381e-2 & 3.500* &  4.84e-4\\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  4.920e-2 &  1.374e-2 & 3.579*  &  3.6e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 5.938e-2 &  3.538e-2  & 1.678* &  1.19e-1 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 4.918e-2 & 1.455e-2 & 3.381** & 7.47e-4 \\
\hline
\end{tabular}

\vspace{5pt}

Table 4: Summary Table for $CD19 \sim MALAT1$ variable parings.  * Approximate normal distribution.   ** Approximate Wald-Z distribution
\end{center}


--------------------------------------------------------------------------

\newpage

<!------------------------------------------------------------------------------>

\small{
\begin{center}
\centering
\begin{tabular}{|p{0.15\textwidth}|p{0.275\textwidth}|c|c|c|c|}
\hline \noalign{\smallskip}
\center{Model \newline Designation} & \center{Model Description} & Estimate & Std. Error & Test Statistic & p-value \\
\hline
\hline
\center{LM} & \center{Linear Model} &  7.884e-1 &  4.92e-2  & 4.002* & <2e-16\\
\hline \noalign{\smallskip}
\center{LM-FE} & \center{Linear Model with  \newline Fixed-Effect Intercept} &  1.31e-1 & 3.42e-2  & 3.818* & 1.42e-4 \\
\hline \noalign{\smallskip}
\center{LMM-RI} & \center{Linear Mixed Model with \newline Random Intercept} &  1.35e-1 & 3.42e-2  & 3.95* & 8.4e-5 \\
\hline \noalign{\smallskip}
\center{LMM-RS} & \center{Linear Mixed Model with \newline Random Slope }  & 1.705e-1  & 7.29e-2 & 2.34*  & 6.7e-2 \\
\hline \noalign{\smallskip}
\center{GEE} & \center{Generalized Estimating Equations } & 7.884e-1 & 4.92e-2 & 4.002** & < 2e-16 \\
\hline
\end{tabular}

\vspace{5pt}

Table 5: Summary Table for $CD34 \sim FBLN$ variable parings.  * Approximate normal distribution.   ** Approximate Wald-Z distribution
\end{center}

}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0 &  -1.7283 &  0.0407 & 20.7401 &  0.0000 \\
\hline
LM-FE & 1.7587 &  0 &  1.8001 & 22.8636  & 1.7587  \\
\hline
LMM-RI & -0.0407 & -1.7683 &  0 & 20.6911 & -0.0407  \\
\hline
LMM-RS & -17.1775 & -18.6090 & -17.1438 & 0 & -17.1775 \\
\hline
GEE & 0.0000 & -1.7283  & 0.0407 & 20.7401  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table 6: Main effect slope Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -83.3841 & -82.8767  & -78.3739 & 0.0000  \\
\hline
LM-FE & 501.8321 & 0   & 3.0534    & 30.1527  & 501.8321 \\
\hline
LM-RI & 484.0000 & -2.9630  & 0    & 26.2963  & 484.0000  \\
\hline
LM-RS & 362.4047 & -23.1672 & -20.8211  & 0   & 362.4047  \\
\hline
GEE & 0.0000   & -83.3841 & -82.8767  & -78.3739 & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 7: Main effect slope Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -5.0859  & -5.5670   & 143.1615  & 0.0000  \\
\hline
LM-FE & 5.3584   & 0   & -0.5069   & 156.1912  & 5.3584  \\
\hline
LMM-RI & 5.8952   & 0.5095   & 0    & 157.4964  & 5.8952  \\
\hline
LMM-RS & -58.8751 & -60.9666 & -61.1645  & 0    & -58.8751  \\
\hline
GEE &  0.0000   & -5.0859  & -5.5670   & 143.1615  & 0        \\
\hline
\end{tabular}

\vspace{5pt}

Table 8: Main effect slope Standar Error Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0   & -30.4878 & -30.4878  & 48.1707   & 0.0000  \\
\hline
LM-FE & 43.8596  & 0   & 0.0000    & 113.1579  & 43.8596  \\
\hline
LM-RI & 43.8596  & 0.0000   & 0    & 113.1579  & 43.8596  \\
\hline
LM-RS & -32.5103 & -53.0864 & -53.0864  & 0    & -32.5103  \\
\hline
GEE &  0.0000   & -30.4878 & -30.4878  & 48.1707   & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 9: Main effect slope Standar Error Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & 3.5197   & 5.8563    & -50.3697  & 0.0000 \\
\hline
LM-FE & -3.4000  & 0   & 2.2571    & -52.0571  & -3.4000 \\
\hline
LMM-RI & -5.5323  & -2.2073  & 0    & -53.1154  & -5.5323  \\
\hline
LM-RS & 101.4899 & 108.5816 & 113.2896  & 0    & 101.4899 \\
\hline
GEE & 0.0000  & 3.5197   & 5.8563    & -50.3697  & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table 10: Main effect slope Test Statistic Percent Change matrix, $CD19 \sim MALAT1$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
Model & LM & LM-FE & LMM-RI & LMM-RS & GEE \\
\hline
\hline
LM & 0  & -4.5977  & -1.2994   & -41.5292  & 0.0000  \\
\hline
LM-FE & 4.8193  & 0   & 3.4573    & -38.7114  & 4.8193  \\
\hline
LM-RI & 1.3165  & -3.3418  & 0    & -40.7595  & 1.3165  \\
\hline
LM-RS & 71.0256 & 63.1624  & 68.8034   & 0    & 71.0256  \\
\hline
GEE & 0.0000  & -4.5977  & -1.2994   & -41.5292  & 0  \\
\hline
\end{tabular}

\vspace{5pt}

Table 11: Main effect slope Test Statistic Percent Change matrix, $CD34 \sim FBLN$ variable pairing
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}

\textbf{\large{CD19 Summaries}}

\begin{tabular}{|c|c|c|c|c|}
\hline
Subject Number & Minimum & Maximum & Average & Median \\
\hline
\hline
5 &  0 & 678 & 36.6724 & 0.0 \\
\hline
6 &  0 & 299 & 36.6860 & 7.5 \\
\hline
7 &  0 & 10 & 2.1250 & 1.0 \\
\hline
9 &  0 & 1052 & 89.4194 & 3.0 \\
\hline
10 & 0 & 158 & 37.5714 & 2.0 \\
\hline
11 & 0 & 339 & 28.3178 & 1.0 \\
\hline
13 & 0 & 629 & 56.0841 & 18.0 \\
\hline
14 & 0 & 251 & 42.2600 & 19.0 \\
\hline
15 & 0 & 148 & 26.6000 & 0.0 \\
\hline
17 & 0 & 982 & 112.3770 & 16.0 \\
\hline
19 & 0 & 665 & 59.3386 & 5.0 \\
\hline
20 & 0 & 287 & 40.1200 & 23.0 \\
\hline
22 & 0 & 380 & 43.4483 & 1.0 \\
\hline
24 & 0 & 282 & 55.0127 & 27.0 \\
\hline
26 & 0 & 1624 & 268.4151 & 110.0 \\
\hline
\end{tabular}

\vspace{5pt}

Table A1: Predictor $CD19$ variable summaries ($CD19 \sim MALAT1$)
\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}

\textbf{\large{MALAT1 Summaries}}

\begin{tabular}{|c|c|c|c|c|}
\hline
Subject Number & Minimum & Maximum & Average & Median \\
\hline
\hline
5  & 67 & 40812 & 10206.3621 & 9195.0 \\
\hline
6  & 757 & 30774 & 11568.2791 & 11689.0 \\
\hline
7  & 441 & 17916 & 6868 & 4039.5 \\
\hline
9  & 311 & 18239 & 5703.9355 & 5983.0 \\
\hline
10 & 1875 & 17160 & 6638.5714 & 6190.0 \\
\hline
11 & 349 & 34082 & 9716.0280 & 8826.0 \\
\hline
13 & 99 & 25572 & 5867.9439 & 4895.0 \\
\hline
14 & 355 & 15740 & 6154.1500 & 5720.5 \\
\hline
15 & 157 & 11923 & 3839.0800 & 3467.0 \\
\hline
17 & 337 & 8342 & 2960.2541 & 2692.0 \\
\hline
19 & 227 & 91961 & 13959.9843 & 10125.0 \\
\hline
20 & 379 & 21736 & 7301.4133 & 6417.0 \\
\hline
22 & 161 & 28429 & 6881.7471 & 5068.0 \\
\hline
24 & 240 & 42792 & 6248.8228 & 5955.0 \\
\hline
26 & 1114 & 32426 & 8463.1698 & 6426.0 \\
\hline
\end{tabular}

\vspace{5pt}

Table A2: Response $MALAT1$ variable summaries ($CD19 \sim MALAT1$)

\end{center}

<!------------------------------------------------------------------------------>

\newpage

<!------------------------------------------------------------------------------>

\begin{center}

\textbf{\large{CD34 Summaries}}


\begin{tabular}{|c|c|c|c|c|}
\hline
Subject Number & Minimum & Maximum & Average & Median \\
\hline
\hline
5  & 0 & 19 & 3.0517 & 1 \\
\hline
6  & 0 & 0 & 0 & 0 \\
\hline
7  & 0 & 0 & 2 & 1 \\
\hline
9  & 0 & 6 & 0.4516 & 0 \\
\hline
10 & 0 & 5 & 0.6667 & 0 \\
\hline
11 & 0 & 7 & 1.2056 & 1 \\
\hline
13 & 0 & 0 & 0 & 0 \\
\hline
14 & 0 & 1 & 0.4000 & 0 \\
\hline
15 & 0 & 0 & 0 & 0 \\
\hline
17 & 0 & 0 & 0 & 0 \\
\hline
19 & 0 & 0 & 0 & 0 \\
\hline
20 & 0 & 2 & 0.1867 & 0 \\
\hline
22 & 0 & 4 & 0.3563 & 0 \\
\hline
24 & 0 & 5 & 0.2911 & 0 \\
\hline
26 & 0 & 0 & 0 & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table A3: Predictor $CD34$ variable summaries ($CD34 \sim FBLN1$)

\end{center}
<!------------------------------------------------------------------------------>


\newpage

<!------------------------------------------------------------------------------>

\begin{center}

\textbf{\large{FBLN1 Summaries}}


\begin{tabular}{|c|c|c|c|c|}
\hline
Subject Number & Minimum & Maximum & Average & Median \\
\hline
\hline
5  & 3 & 41 & 19.3448 & 18 \\
\hline
6  & 0 & 0 & 0 & 0 \\
\hline
7  & 0 & 16 & 4.2500 & 3 \\
\hline
9  & 0 & 8 & 1.8710 & 1 \\
\hline
10 & 0 & 30 & 11.9524 & 10 \\
\hline
11 & 0 & 8 & 1.5140 & 1 \\
\hline
13 & 0 & 1 & 0.0093 & 0 \\
\hline
14 & 0 & 5 & 0.5700 & 0 \\
\hline
15 & 0 & 1 & 0.0400 & 0 \\
\hline
17 & 0 & 3 & 0.0246 & 0 \\
\hline
19 & 0 & 2 & 0.0157 & 0 \\
\hline
20 & 0 & 9 & 2.5867 & 2 \\
\hline
22 & 0 & 11 & 0.9885 & 0 \\
\hline
24 & 0 & 4 & 0.4557 & 0 \\
\hline
26 & 0 & 0 & 0 & 0 \\
\hline
\end{tabular}

\vspace{5pt}

Table A4: Response $FBLN1$ variable summaries ($CD34 \sim FBLN1$)
\end{center}


<!------------------------------------------------------------------------------>

\newpage
